---
title: "Data-Driven Coupon Acceptance Analysis"
author: "Group 2: 
Ayse Ciftci (fXXX),
Christos Kapotis (fXXX),
Ioannis Koutsogiannis (fXXX),
Lazar Nenovski (fXXX),
Dallandyshe Shabani (fXXX)
"
date: today
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-tools: true
  pdf:
    documentclass: scrartcl
    pdf-engine: xelatex
    mainfont: Source Sans Pro
    monofont: Fira Code
    fontsize: 11pt
    geometry:
      - top=2.2cm
      - bottom=2.2cm
      - left=2.2cm
      - right=2.2cm
    number-sections: true
    colorlinks: true
    fig-width: 6.5
    fig-height: 4
    fig-align: center
 
    # Stylistic options for the LaTeX render
    header-includes:
      - \usepackage{fontspec}
      - \usepackage{microtype}
      - \usepackage{booktabs}
      - \usepackage{siunitx}
      - \usepackage{caption}
      - \setlength{\parindent}{0pt}
      - \setlength{\parskip}{6pt}
      - \setkomafont{section}{\normalfont\Large\bfseries}
      - \setkomafont{subsection}{\normalfont\large\bfseries}
      - \usepackage{xcolor}
      - \definecolor{inlinecode}{HTML}{7d12ba}
      - \definecolor{inlinebg}{HTML}{f8f9fa}
      - \let\oldtexttt\texttt
      - \renewcommand{\texttt}[1]{\colorbox{
            inlinebg}{\oldtexttt{\textcolor{inlinecode}{#1}}}}
      - \definecolor{codebg}{HTML}{F5F5F5}
      - \definecolor{outbg}{HTML}{E8E8E8}
    tbl-colwidths: auto

highlight-style: github

execute:
  warning: false
  message: false
  # echo: true
---

# Introduction

\<Introduction goes here \>

# Data Loading & Preprocessing

## Load Libraries

Command to load the R libraries we will use for the analysis:

```{r, library-setup}
library(tidyverse)
library(caret)
library(randomForest)
library(pROC)
library(ggplot2)
library(janitor)
library(rpart)
library(rpart.plot)
```

## Load Data

Here, we load our csv data set into an dataframe and perform an initial 
inspection:

```{r, load-csv}
# for installing packages also needed 
# ("  ,dependencies = TRUE, method = "wininet" ")
df <- read.csv("./dataset/coupons.csv")
glimpse(df)
# summary(df)
```

## Process Data

### Data Cleaning and Categorization of Features

#### Initial cleaning
We will use `clean_names` in order to avoid any inconsistencies in the string
values. Afterwards we will automatically convert all character features into
factors, because all of them describe a category. This will allow us to
reason easier when performing data cleaning and exploratory analysis. We will
also set any empty string as NA, as that's how it's represented in the data set.

```{r, autoset-chr-fact}
df <- clean_names(df)
df <- df %>% mutate_if(is.character, list(~na_if(.,"")))
df <- df %>%
  mutate(across(where(is.character), as.factor))
summary(df)

```
From this, we can observe that in our data the average acceptance rate is 
`0.5684`. Therefore we will accept this as a benchmark of how a naive model
will perform on this data set.
```{r, find-target-mean}
naive_target <- mean(df$y)
```

```{r, find-empty-car}
empty_car <- sum(is.na(df$car))
empty_car_per <- empty_car / nrow(df)
sprintf("Empty car values: %d Percentage from total: %f",
        empty_car,
        empty_car_per)
print("Summary of to_coupon_geq5min")
print(summary(df$to_coupon_geq5min))
rm(empty_car, empty_car_per)
```
From the initial summary we can make two observations. The column `car` has a 
total of `12576` empty entries, which is about **99%** of our dataset. Another 
outlier is the `to_coupon_geq5min` column, which only has values of `1`. 
Therefore we will opt to drop the `car` column, as it doesn't constitute any
meaningful information and we will note that in a future survey it could be
tested to hand out the coupons in a distance below 15 minutes from the venue.
```{r, remove-car}
df <- df %>% select(-car)
```
We will investigate and validate the 15 min and 25 min distance columns, it 
should follow, that `geq15min` is a proper subset of `geq25min`. We already know
that both are a proper subset of `geq5min`, as each value is marked as `1`.
```{r, check-to_coupon}
sum(df$to_coupon_geq25min == 1 & df$to_coupon_geq15min == 0)
```
That seems to indeed be the case, therefore we can conclude that the data about
the distance is valid. However, it contains in itself a large amount of 
redundant information, since it follows an ascending inequality. Since all three
columns are a Yes/No categorical value, and since all elements that fit the
category of `geq25` naturally fit into both `geq15` and `geq5` we will fold 
those three columns into one categorical value, called `distance_toCoupon`, as
an ordered factor as distance increases:
```{r, fold-to_coupon}
df$distance_toCoupon <- case_when(
    df$to_coupon_geq25min == 1 ~ "geq25",
    df$to_coupon_geq15min == 1 ~ "geq15",
    df$to_coupon_geq5min == 1 ~ "geq5",
    TRUE ~ NA) # Should not occur in the current dataset.

df$distance_toCoupon <- factor(df$distance_toCoupon,
                               levels = c("geq5", "geq15", "geq25"),
                               ordered = TRUE)
df <- df %>% select(-to_coupon_geq5min, -to_coupon_geq15min, -to_coupon_geq25min)
```

We can investigate and verify similar information duplication in the columns
`direction_same` and `direction_opp`, given that there are two different 
possibilities for a direction. We will first verify the consistency of the
data. From the initial summary, it seems that the two columns are a compliment
from each other:
```{r, check-direction}
sprintf("Opposite direction : %d", sum(df$direction_opp))
sprintf("Same direction: %d", sum(df$direction_same))
sprintf("Total rows in any directions: %d", 
        sum(df$direction_opp) + sum(df$direction_same))
sprintf("Total rows: %d", nrow(df))
both_1 <- nrow(df[df$direction_opp == 1 & df$direction_same == 1, ])
both_0 <- nrow(df[df$direction_opp == 0 & df$direction_same == 0, ])
sprintf("Both in same and opposite distance: %d", both_1)
sprintf("Neither in the same and opposite distance: %d", both_0)
rm(both_0, both_1)
```
Since those columns are a perfect compliment of each other, we will drop the
column and only use `direction_op`, since it is the more common occurrence and
we will use it as a reference value.
```{r, clean-direction}
df <- df %>% select(-direction_same)
df$direction_opp <- factor(df$direction_opp,
                           levels = c(0, 1),
                           labels = c("No", "Yes"))
df$direction_opp <- relevel(df$direction_opp, ref="Yes")
```

#### Categorization

We will convert our variable encoded as numerical 1/0 into a factor, since it
also represents a categorical value and it will make it easier to analyze 
further. We will also set the reference for all categorical variables.
For Yes/No type features or classification, we will use 'No' or 'never' neutral 
state as a reference. This applies to the columns `has_children`, `Bar`, 
`CoffeeHouse`, `CarryAway`, `RestaurantLessThen20`, `Restaurant20To50`, 
`toCoupon_*`, `direction_*`. Additionally for our target variable, `Y`, we will
use "Rejected" as a reference.

```{r, categorize-binary}
#Shorthand
t_01 <- c(0, 1)
t_NY <- c("No", "Yes")

# Shorthand for relabeling 0/1 data to a set of labels
binary_to_factor <- function(x, lbls = t_NY) {
  x <- factor(x, levels = c(0,1), labels = lbls)
  relevel(x, lbls[1]) # redundant
}

df$y <- binary_to_factor(df$y, lbls = c("Rejected", "Accepted")) # Target
df$y <- relevel(df$y, ref = "Rejected")

df$has_children <- binary_to_factor(df$has_children, 
                                      lbls = c("No children", "Has children"))
cols_yn <- c()
#cols_yn <- c("direction_opp",
#            "direction_same",
##            "to_coupon_geq5min",
#            "to_coupon_geq15min",
#            "to_coupon_geq25min")
#
#df <- df %>%
#  mutate(across(all_of(cols_yn), binary_to_factor))

rm(t_01, t_NY, cols_yn, binary_to_factor)
#df <- distinct(df)
#summary(df)
```
For the categories that represent ordered data, we will relevel them, as to 
indicate the ordering, from lowest to highest. For the unordered type of
categories, we will use the most common value as a reference. 

```{r, categorize-1-8}
factor_0to8 <- function(x) {
  factor(x, levels = c("never", "less1", "1~3", "4~8", "gt8"), ordered = TRUE)
}
cols_0to8 <- c("bar", 
               "coffee_house", 
               "carry_away", 
               "restaurant_less_than20", 
               "restaurant20to50")
df <- df %>%
  mutate(across(all_of(cols_0to8), factor_0to8))
rm(cols_0to8, factor_0to8)

df$education <- factor(
                  df$education,
                  levels = c("Some High School",
                             "High School Graduate",
                             "Some college - no degree",
                             "Associates degree",
                             "Bachelors degree",
                             "Graduate degree (Masters or Doctorate)"),
                  ordered = TRUE)

df$age <- factor(df$age,
                 levels = c("below21", 
                            "21",
                            "26",
                            "31",
                            "36",
                            "41",
                            "46",
                            "50plus"),
                 ordered = TRUE)
df$income <- factor(df$income,                 
                    levels = c("Less than $12500",
                               "$12500 - $24999",
                               "$25000 - $37499",
                               "$37500 - $49999",
                               "$50000 - $62499",
                               "$62500 - $74999",
                               "$75000 - $87499",
                               "$87500 - $99999",
                               "$100000 or More"),
                    ordered = TRUE)

df$temperature <- factor(df$temperature, ordered = TRUE)
df <- df %>%
  mutate(
    destination = relevel(destination, ref = "No Urgent Place"),
    passanger = relevel(passanger, ref = "Alone"),
    weather = relevel(weather, ref = "Sunny"),
    time = relevel(time, ref = "10AM"),
    coupon = relevel(coupon, ref = "Coffee House"),
    marital_status = relevel(marital_status, ref = "Single"),
    occupation = relevel(occupation, ref = "Unemployed"),
  )
summary(df)
```
## Split and treatment of NA
Since our objective is to determine if a customer will accept or reject a coupon
for a specific venue, we will investigate several cases: First for a general
acceptance, using the entire dataset and second, that investigates the factors
for each specific coupon type. Therefore we will split the dataset several ways.

### Overall model data preparation
We will investigate and prepare the entire dataset for analysis and modelling.
While most of the rows are complete, there are several that have NA values.
```{r, check-noncomplete}
sprintf("Noncomplete records in dataset: %f", sum(!complete.cases(df)) / nrow(df))
df_full_complete <- na.omit(df)
```
Sine the number of noncomplete records is less then 5%, we will drop those rows.


```{r, check-missing}
# Check missing values
colSums(is.na(df))

table(df$car, useNA = "always")
```

### Transform the expiration values in hours

Suppose your dataframe is called df and your column is 'expiration'
Function to convert any single value to hours

```{r, convert_hours}
convert_to_hours <- function(x) {
  # Extract all number-unit pairs (works for compound like "1d 2h 30m")
  matches <- str_match_all(x, "(\\d+\\.?\\d*)\\s*([a-zA-Z]+)")[[1]]
  
  if(nrow(matches) == 0) return(NA_real_)  # fallback if no match
  total_hours <- 0
  
  for(i in 1:nrow(matches)) {
    num <- as.numeric(matches[i,2])
    unit <- tolower(matches[i,3])
    
    total_hours <- total_hours + switch(unit,
                                        "m" = num / 60,
                                        "h" = num,
                                        "d" = num * 24,
                                        "w" = num * 24 * 7,
                                        NA_real_)
  }
  
  return(total_hours)
}
```

Apply function to your whole column dynamically

```{r, apply-hoursconvert}
df <- df %>%
  mutate(
    expiration_original = expiration,   # backup column
    expiration = vapply(expiration, convert_to_hours, numeric(1))
  )

df_bckup <- df
```

### Transform the expiration values in hours

transform time to time hour numeric from AM/PM

```{r, parse-datetime}
df <- df %>%
  mutate(time_original = time,
         time = hour(parse_date_time(time, orders = "I!p")))
```
### replace "" with NAs and remove them 

```{r, freq-clean}
sapply(df[, c("bar",
              "coffee_house",
              "carry_away",
              "restaurant_less_than20",
              "restaurant20to50")],
       function(x) table(x, useNA = "always"))

freq_cols <- c("bar",
               "coffee_house",
               "carry_away",
               "restaurant_less_than20",
               "restaurant20to50")

df[freq_cols] <- lapply(df[freq_cols], function(x) {
  x[x == ""] <- NA
  x
})

df[freq_cols] <- lapply(df[freq_cols], droplevels) # drop unused levels pair with the above one next command

sapply(df[freq_cols], function(x) table(x, useNA = "always"))

# drop the cols we kept for backup
df$expiration_original <- NULL
df$time_original  <- NULL

# i will create two y y_numeric and y_factor the numeric will be for plotting and trends and y_factor will be the target variable for modeling (classification).
df <- df %>%
  mutate(
    y_numeric = (as.numeric(y) - 1),
    y_factor = factor(y, levels = c(0,1))
  )
```

-------------------------------------------------


# Exploratory Data Analysis

## Coupon Type vs Acceptance Rate (Core Business Insight) {#fig-coupon}

```{r, coupon-type-vs-acceptance-rate}
df %>%
  group_by(coupon) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = reorder(coupon, accept_rate), y = accept_rate)) +
  geom_col(fill = "#2c7fb8") +
  coord_flip() +
  labs(title = "Coupon Acceptance Rate by Coupon Type",
       x = "Coupon Type",
       y = "Acceptance Rate") +
  theme_minimal()
```
This visualization presents the acceptance rate for each coupon category. The 
results show that **Coffee House and Bar coupons have relatively higher 
acceptance rates compared to other categories**. One possible explanation is 
that these types of places are **easy for customers to visit as part of their 
daily routine**, for example grabbing a coffee or meeting friends for a drink. 
In contrast, **Restaurant(20–50) coupons have lower acceptance rates**, possibly
because going to a restaurant usually requires **more time and planning compared 
to a quick coffee or bar visit**.

## Acceptance Rate by Time of Day (Behavioral Timing Insight)  {#fig-time}
```{r, acceptance-rate-by-time-of-day}
df %>%
  group_by(time) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = time, y = accept_rate, group = 1)) +
  geom_line(color = "#e41a1c", size = 1.2) +
  geom_point(size = 3, color = "#e41a1c") +
  labs(title = "Coupon Acceptance Rate by Time of Day",
       x = "Time of Day",
       y = "Acceptance Rate") +
  theme_minimal()
```

This chart examines how coupon acceptance changes depending on the **time of 
day**. The results suggest that **acceptance is slightly higher during the 
morning and afternoon periods** compared to later hours. This may be because 
customers are more likely to use coupons for **daily activities such as coffee 
or quick meals during these times**. Overall, the visualization indicates that 
**timing may influence how likely customers are to accept a coupon**.

## Income vs Acceptance (Customer Affordability Pattern)

```{r, income-vs-acceptance}
df %>%
  group_by(income) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = income, y = accept_rate)) +
  geom_col(fill = "#4daf4a") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Acceptance Rate by Income Level",
       x = "Income Level",
       y = "Acceptance Rate")
```

This visualization explores the relationship between **income level and coupon 
acceptance**. The results show that **customers in lower and middle income 
groups tend to accept coupons more often**. This pattern may suggest that 
discounts and promotions are more attractive to customers who are more careful 
about spending. In contrast, **higher income groups show slightly lower 
acceptance rates**, which may indicate that price incentives are less important 
for them.

## Coffee House Visits vs Acceptance (Behavioral Engagement Signal)

```{r, coffe-vists-vs-acceptance}
df %>%
  group_by(coffee_house) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = coffee_house, y = accept_rate)) +
  geom_col(fill = "#984ea3") +
  labs(title = "Acceptance Rate by Coffee Visit Frequency",
       x = "Monthly Coffee Visits",
       y = "Acceptance Rate") +
  theme_minimal()
```

This chart analyzes the relationship between **how often customers visit coffee 
houses and their likelihood of accepting coupons**. Customers who visit coffee 
houses more frequently show **higher acceptance rates for these types of 
coupons**. This suggests that people who already have a habit of visiting coffee
shops are more likely to take advantage of related offers. Therefore, **past 
behavior appears to be an important factor in predicting coupon acceptance.**

## Heatmap – Coupon Type × Time (Segmentation Insight)

```{r, coupontype-by-time}
df %>%
  group_by(coupon, time) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = coupon, y = time, fill = accept_rate)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "darkgreen") +
  labs(title = "Acceptance Rate: Coupon Type by Time of Day",
       x = "Coupon Type",
       y = "Time of Day",
       fill = "Acceptance Rate") +
  theme_minimal()
```

The heatmap combines **coupon type and time of day** to show how acceptance 
rates vary across different situations. Some combinations show **higher 
acceptance levels, such as Coffee House coupons during morning hours**. Other 
combinations appear less effective, indicating that **not all offers perform the
same at different times of the day**. This suggests that **both the type of 
coupon and the time it is offered may influence customer response**.

## Regression analysis

We wil now run the entire dataset through a general linear regression 
model in order to gain some further insights from the data

```{r, data-summary-before-reg}
df_no_extra_col <- df %>% select(-y_numeric, -y_factor)
summary(df_no_extra_col)
```
```{r, data-summary-reg}
reg_analysis <- glm(y ~ ., data=df_no_extra_col, family=binomial)
summary(reg_analysis)
```

# Predictive Modeling

We will use 3 models:  Logistic Regression, Decision Tree, Random Forest
Remove y_numeric
Keep y_factor as target
Remove irrelevant columns (IDs etc.)

## Create Test and Training dataset

Remove rows with NA

Train/Test Split (70% / 30%)

positive class -> 1 not 0
```{r, data-split}
df_clean <- df %>%
  select(-y_numeric, -y_factor)

df_nonnull <- na.omit(df_clean)

set.seed(123)

train_index <- createDataPartition(df_nonnull$y, p = 0.7, list = FALSE)

train_data <- df_nonnull[train_index, ]
test_data  <- df_nonnull[-train_index, ]

#train_data$y_factor <- relevel(train_data$y_factor, ref = "1")
#test_data$y_factor  <- relevel(test_data$y_factor, ref = "1")

```

## Logistic Regression

```{r, model-logreg}
model_log <- glm(y ~ ., 
                 data = train_data,
                 family = binomial)

prob_log <- predict(model_log, test_data, type = "response")

# Reverse threshold
pred_log <- ifelse(prob_log > 0.5, 1, 0)

pred_log <- factor(pred_log, levels = c(0,1), labels=c("Rejected", "Accepted"))

confusionMatrix(pred_log, test_data$y)
```

## Decision Tree

```{r, model-dectree}
model_tree <- rpart(y ~ ., 
                    data = train_data,
                    method = "class")

rpart.plot(model_tree)

pred_tree <- predict(model_tree, test_data, type = "class")

confusionMatrix(pred_tree, test_data$y)
```

## Random Forest

```{r, model-rforest}
# I'm not sure where this came from, but it seems to not work
#varImpPlot(model_rf) # check for near zero importance variables if yes then delete them and retrain model

model_rf <- randomForest(y ~ ., 
                         data = train_data,
                         ntree = 200)
pred_rf <- predict(model_rf, test_data)

confusionMatrix(pred_rf, test_data$y)

rf_probs <- predict(model_rf, test_data, type = "prob")[,2]
roc_rf <- roc(test_data$y, rf_probs)

auc(roc_rf)
plot(roc_rf)
```

# Model Evaluation

## Calculate Logistic Regression probabilities (probability of class 1)

```{r, eval-logreg}
log_probs <- predict(model_log, test_data, type = "response")
roc_log  <- roc(test_data$y, log_probs)
```

## Decision Tree probabilities

```{r, eval-dectree}
tree_probs <- predict(model_tree, test_data, type = "prob")[,2]
roc_tree <- roc(test_data$y, tree_probs)

```

## Random Forest probabilities

```{r, eval-rforest}
rf_probs <- predict(model_rf, test_data, type = "prob")[,2]
roc_rf   <- roc(test_data$y, rf_probs)
```


## Plot All ROC Curves on One Graph

```{r, plot-all-roc}
plot(roc_log, 
     col = "blue", 
     lwd = 2,
     main = "ROC Curve Comparison")

plot(roc_tree, 
     col = "green", 
     lwd = 2,
     add = TRUE)

plot(roc_rf, 
     col = "red", 
     lwd = 2,
     add = TRUE)

legend("bottomright",
       legend = c(
         paste("Logistic Regression (AUC =", round(auc(roc_log),3),")"),
         paste("Decision Tree (AUC =", round(auc(roc_tree),3),")"),
         paste("Random Forest (AUC =", round(auc(roc_rf),3),")")
       ),
       col = c("blue", "green", "red"),
       lwd = 2)
```

# Data-Driven Marketing Campaign

Based on the insights gathered through exploratory data analysis and predictive
modeling, we propose a comprehensive, data-driven marketing campaign aimed at
maximizing coupon acceptance rates while minimizing unnecessary promotional 
costs.

## Key Customer Segments

The analysis revealed that coupon acceptance is not uniform across all 
customers. We identified four distinct customer segments that differ 
significantly in their likelihood to accept a coupon:

| Segment | Profile | Best Coupon Type | Expected Acceptance |
|---|---|---|---|
| *Coffee Addict* | Visits coffee house 4+ times/month | Coffee House | ~65% |
| *Lunch Crowd* | Drives at midday, no urgent destination | Carry out / Restaurant(<$20) | ~70% |
| *Night Out* | Drives in the evening with friends | Bar / Restaurant($20–$50)| ~55% |
| *Family Driver* | Has children, drives with kids | Carry out & Take away | ~70% |

These segments were derived from the behavioral variables in the dataset
(coffee_house, bar, carry_away, passanger, time) and validated through
the decision tree splits identified in Section 4.3.

## Optimal Targeting Strategy

### Timing

The acceptance rate analysis by time of day (Section 3.2) shows that coupon
acceptance peaks at *2PM (~67%)* and is lowest at *10PM (~52%)*. We therefore 
recommend the following send windows:

| Time Window | Recommended Coupon | Rationale |
|---|---|---|
| 10:00 – 12:00 | Coffee House | Morning coffee break — high demand |
| 12:00 – 15:00 | Carry out / Restaurant(<$20) | Lunch peak — highest acceptance |
| 18:00 – 21:00 | Bar / Restaurant($20–$50) | Evening outing with friends |
| Avoid 22:00+ | All types | Lowest acceptance — budget waste |

### Coupon Expiration

Expiration time plays a significant role in acceptance decisions:

- *2-hour coupons* are most effective during midday hours when the customer is 
  near the venue and can act immediately. They create a sense of urgency.
- *1-day coupons* are better suited for Bar and Restaurant($20–$50) offers, 
  where customers need time to plan an evening outing.

*Recommendation:* Send 2-hour coupons at lunchtime for fast food/carry out, and 
1-day coupons in the morning for evening dining or bar visits.

### Predictive Scoring Threshold

The Random Forest model generates a probability score (0–1) for each customer,
representing their likelihood of accepting a coupon. We propose a
threshold-based targeting system to optimize campaign ROI:

| Score Range | Action | Rationale |
|---|---|---|
| Score > 0.70 | Send coupon immediately | High acceptance probability |
| Score 0.50 – 0.70 | Send with A/B testing | Moderate probability — worth testing |
| Score < 0.50 | Do not send | Low probability — cost saving |

This approach reduces unnecessary coupon distribution while focusing budget
on customers most likely to respond positively.

## Budget Allocation

Based on the coupon-type acceptance rates identified in Section 3.1,
we recommend the following budget distribution:

| Coupon Type | Acceptance Rate | Priority | Recommended Budget Share |
|---|---|---|---|
| Carry out & Take away | ~70% | ★★★ High | 30% |
| Restaurant(<$20) | ~70% | ★★★ High | 30% |
| Coffee House | ~65% | ★★ Medium | 20% |
| Restaurant($20–$50) | ~58% | ★★ Medium | 15% |
| Bar | ~41% | ★ Low | 5% |

Bar coupons should only be distributed to highly targeted segments
(frequent bar visitors, evening drivers with friends) to avoid
low-return spend.

## Marketing Channels & Tactics

| Channel | Target Segment | Tactic |
|---|---|---|
| Mobile push notification | All segments | Send coupon during identified high-acceptance time windows |
| In-app GPS trigger | Lunch Crowd, Coffee Addict | Trigger coupon when driver is within 15 min of venue |
| Email (morning send) | Night Out segment | Send 1-day Bar/Restaurant coupons before 10AM |
| Loyalty program integration | Frequent visitors (4~8, gt8) | Offer escalating rewards for repeat coupon usage |

## Key Performance Indicators (KPIs)

To evaluate the success of the campaign, we propose monitoring the following 
KPIs on a monthly basis:

| KPI | Definition | Target |
|---|---|---|
| Coupon Redemption Rate | % of sent coupons that were used | > 60% |
| Cost per Redemption | Campaign cost / number of redemptions | Reduction vs. baseline |
| Revenue Lift | Revenue increase vs. period without coupons | > 15% |
| Model Accuracy | Ongoing accuracy of the Random Forest model | Maintain > 75% |
| Segment Response Rate | Acceptance rate per customer segment | Improvement per segment |

## Limitations & Future Improvements

We acknowledge the following limitations of this analysis:

- The dataset is based on a *survey*, not real transaction data. Actual
  behavior in real-world conditions may differ from survey responses.
- The model lacks *real-time GPS data*, which would significantly improve
  location-based targeting precision.
- The model requires *periodic retraining* as consumer habits evolve over time.
- The analysis does not account for *seasonal variation* (e.g., different 
  behavior in winter vs. summer).

To further improve campaign effectiveness, we recommend:

- Implementing *cross-validation* for more robust model evaluation.
- Applying *hyperparameter tuning* to further optimize the Random Forest model.
- Integrating *real-time location data* for dynamic, proximity-based coupon 
  delivery.
- Establishing an *A/B testing framework* to continuously compare targeting 
  strategies.

------------------------------------------------------------------------
