---
title: "Data-Driven Coupon Acceptance Analysis"
author: "Group 2:"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-tools: true
execute:
  warning: false
  message: false
  echo: true
---

# Introduction

\<Introduction goes here \>

# Data Loading & Preprocessing

## Load Libraries

Command to load the R libraries we will use for the analysis:

```{r}
library(tidyverse)
library(caret)
library(randomForest)
library(pROC)
library(ggplot2)
library(janitor)
library(rpart)
library(rpart.plot)
```

## Load Data

Here, we load our csv data set into an dataframe and perform an initial 
inspection:

```{r}
### for installing packages also needed ("  ,dependencies = TRUE, method = "wininet" ")
df <- read.csv("./dataset/coupons.csv")

glimpse(df)
summary(df)
```

## Process Data

### Data Cleaning and Categorization of Features

### Initial cleaning
We will use `clean_names` in order to avoid any inconsistencies in the string
values. Afterwards we will automatically convert all character features into
factors, because all of them describe a category. This will allow us to
reason easier when performing data cleaning and exploratory analysis. We will
also set any empty string as NA, as that's how it's represented in the data set.

```{r}
df <- clean_names(df)

df <- df %>% mutate_if(is.character, list(~na_if(.,""))) 

df <- df %>%
  mutate(across(where(is.character), as.factor))
summary(df)

```
From this, we can observe that in our data the average acceptance rate is 
`0.5684`. Therefore we will accept this as a benchmark of how a naive model
will perform on this data set.
```{R}
naive_target <- mean(df$y)
```

```{R}
empty_car <- sum(df$car == "")
empty_car_per <- empty_car / nrow(df)
print(sprintf("Empty car values: %d Percentage from total: %f",
              empty_car,
              empty_car_per))
print("Summary of to_coupon_geq5min")
print(summary(df$to_coupon_geq5min))
rm(empty_car, empty_car_per)
```
From the initial summary we can make two observations. The column `car` has a 
total of `12576` empty entries, which is about **99%** of our dataset. Another 
outlier is the `to_coupon_geq5min` column, which only has values of `1`. 
Therefore we will opt to drop those two column, as they don't constitute any
meaningful information and we will note that in a future survey it could be
tested to hand out the coupons in a distance below 15 minutes from the venue.
```{R}
df <- df %>% select(-car, -to_coupon_geq5min)
```
We will quickly validate the 15 min and 25 min distance columns, it should 
follow, that `geq15min` is a proper subset of `geq25min`
```{R}
sum(df$to_coupon_geq25min == 1 & df$to_coupon_geq15min == 0)
```
That seems to indeed be the case, therefore we will not take any further 
actions.


**CHECK ME**
Although here we drop it, it might be better to consider if we should pivot those
three columns into one about distance.
**CHECK ME**

### Categorization

We will convert our variable encoded as numerical 1/0 into a factor, since it
also represents a categorical value and it will make it easier to analyze 
further. We will also set the reference for all categorical variables.
For Yes/No type features or classification, we will use 'No' or 'never' neutral 
state as a reference. This applies to the columns `has_children`, `Bar`, 
`CoffeeHouse`, `CarryAway`, `RestaurantLessThen20`, `Restaurant20To50`, 
`toCoupon_*`, `direction_*`. Additionally for our target variable, `Y`, we will
use "Rejected" as a reference.

```{R}
#Shorthand
t_01 <- c(0, 1)
t_NY <- c("No", "Yes")

# Shorthand for relabeling 0/1 data to a set of labels
binary_to_factor <- function(x, lbls = t_NY) {
  x <- factor(x, levels = c(0,1), labels = lbls)
  relevel(x, lbls[1]) # redundant
}

df$y <- binary_to_factor(df$y, lbls = c("Rejected", "Accepted")) # Target

df$has_children <- binary_to_factor(df$has_children, 
                                      lbls = c("No children", "Has children"))
cols_yn = c("direction_opp",
            "direction_same",
#            "to_coupon_geq5min",
            "to_coupon_geq15min",
            "to_coupon_geq25min")

df <- df %>%
  mutate(across(all_of(cols_yn), binary_to_factor))

rm(t_01, t_NY, cols_yn, binary_to_factor)
#df <- distinct(df)
#summary(df)
```
For the categories that represent ordered data, we will relevel them, as to 
indicate the ordering, from lowest to highest. For the unordered type of
categories, we will use the most common value as a reference. 

```{R}
factor_0to8 <- function(x) {
  factor(x, levels = c("never", "less1", "1~3", "4~8", "gt8"), ordered = TRUE)
}
cols_0to8 <- c("bar", 
               "coffee_house", 
               "carry_away", 
               "restaurant_less_than20", 
               "restaurant20to50")

df <- df %>%
  mutate(across(all_of(cols_0to8), factor_0to8))

rm(cols_0to8, factor_0to8)
```

```{R}
# Check missing values
colSums(is.na(df))

table(df$y)
prop.table(table(df$y))
#they return 0 1 and sum how many zeros and ones
#and the prop the same but perc in based total count results
#check if the dataset is imbalanced.
#This is extremely important for modeling quality.

table(df$car, useNA = "always")
```

### Transform the expiration values in hours

Suppose your dataframe is called df and your column is 'expiration'
Function to convert any single value to hours

```{R}
convert_to_hours <- function(x) {
  # Extract all number-unit pairs (works for compound like "1d 2h 30m")
  matches <- str_match_all(x, "(\\d+\\.?\\d*)\\s*([a-zA-Z]+)")[[1]]
  
  if(nrow(matches) == 0) return(NA_real_)  # fallback if no match
  total_hours <- 0
  
  for(i in 1:nrow(matches)) {
    num <- as.numeric(matches[i,2])
    unit <- tolower(matches[i,3])
    
    total_hours <- total_hours + switch(unit,
                                        "m" = num / 60,
                                        "h" = num,
                                        "d" = num * 24,
                                        "w" = num * 24 * 7,
                                        NA_real_)
  }
  
  return(total_hours)
}
```

Apply function to your whole column dynamically

Check result

```{R}

df <- df %>%
  mutate(
    expiration_original = expiration,   # backup column
    expiration = vapply(expiration, convert_to_hours, numeric(1))
  )


head(df)

df_bckup <- df
```

### Transform the expiration values in hours

transform time to time hour numeric from AM/PM

```{R}
df <- df %>%
  mutate(time_original = time,
         time = hour(parse_date_time(time, orders = "I!p")))
```

### Age

Age was treated as an ordinal variable since the dataset provides age groups rather than continuous values.

```{R}
# check if a column is factor or not #
class(df$age)

df$age <- factor(df$age,
                 levels = c("below21","21","26","31","36","41","46","50plus"),
                 ordered = TRUE) #Age is numeric — not categorical. Otherwise the model treats age as categories instead of a continuous variable.
table(df$age, useNA = "always")

```


### replace "" with NAs and remove them 

```{R}
sapply(df[, c("bar",
              "coffee_house",
              "carry_away",
              "restaurant_less_than20",
              "restaurant20to50")],
       function(x) table(x, useNA = "always"))

freq_cols <- c("bar",
               "coffee_house",
               "carry_away",
               "restaurant_less_than20",
               "restaurant20to50")

df[freq_cols] <- lapply(df[freq_cols], function(x) {
  x[x == ""] <- NA
  x
})

df[freq_cols] <- lapply(df[freq_cols], droplevels) # drop unused levels pair with the above one next command

sapply(df[freq_cols], function(x) table(x, useNA = "always"))

colSums(is.na(df)) # count NAs per column 100 - 200 per column so small percentage based ~5% so we can delete the rows

df <- na.omit(df) #delete rows "Observations with missing frequency values were removed due to their low proportion in the dataset (<5%)."


# We should convert frequency variables to ordered factors:
df[freq_cols] <- lapply(df[freq_cols], function(x) {
  factor(x,
         levels = c("never","less1","1~3","4~8","gt8"),
         ordered = TRUE)
})

str(df$coffee_house)

# drop the cols we kept for backup
df$expiration_original <- NULL
df$time_original  <- NULL

# i will create two y y_numeric and y_factor the numeric will be for plotting and trends and y_factor will be the target variable for modeling (classification).
df <- df %>%
  mutate(
    y_numeric = as.numeric(as.character(y)),
    y_factor = factor(y, levels = c(0,1))
  )
```

-------------------------------------------------


# Exploratory Data Analysis

## Coupon Type vs Acceptance Rate (Core Business Insight) {#fig-coupon}

```{r}
df %>%
  group_by(coupon) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = reorder(coupon, accept_rate), y = accept_rate)) +
  geom_col(fill = "#2c7fb8") +
  coord_flip() +
  labs(title = "Coupon Acceptance Rate by Coupon Type",
       x = "Coupon Type",
       y = "Acceptance Rate") +
  theme_minimal()
```

## Acceptance Rate by Time of Day (Behavioral Timing Insight)  {#fig-time}
```{R}
df %>%
  group_by(time) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = time, y = accept_rate, group = 1)) +
  geom_line(color = "#e41a1c", size = 1.2) +
  geom_point(size = 3, color = "#e41a1c") +
  labs(title = "Coupon Acceptance Rate by Time of Day",
       x = "Time of Day",
       y = "Acceptance Rate") +
  theme_minimal()
```

## Income vs Acceptance (Customer Affordability Pattern)

```{R}
df %>%
  group_by(income) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = income, y = accept_rate)) +
  geom_col(fill = "#4daf4a") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Acceptance Rate by Income Level",
       x = "Income Level",
       y = "Acceptance Rate")
```

## Coffee House Visits vs Acceptance (Behavioral Engagement Signal)
```{R}
df %>%
  group_by(coffee_house) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = coffee_house, y = accept_rate)) +
  geom_col(fill = "#984ea3") +
  labs(title = "Acceptance Rate by Coffee Visit Frequency",
       x = "Monthly Coffee Visits",
       y = "Acceptance Rate") +
  theme_minimal()
```
## Heatmap – Coupon Type × Time (Segmentation Insight)
```{R}
df %>%
  group_by(coupon, time) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = coupon, y = time, fill = accept_rate)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "darkgreen") +
  labs(title = "Acceptance Rate: Coupon Type by Time of Day",
       x = "Coupon Type",
       y = "Time of Day",
       fill = "Acceptance Rate") +
  theme_minimal()
```

# Predictive Modeling

We will use 3 models:  Logistic Regression, Decision Tree, Random Forest
Remove y_numeric
Keep y_factor as target
Remove irrelevant columns (IDs etc.)

## Create Test and Training dataset

Remove rows with NA

Train/Test Split (70% / 30%)

positive class -> 1 not 0
```{R}
#df_clean <- df %>%
#  select(-y, -y_numeric)

#df_model <- na.omit(df_clean)

set.seed(123)

train_index <- createDataPartition(df_nonnull$y, p = 0.7, list = FALSE)

# train_data <- df_nonnull[train_index, ]
test_data  <- df_nonnull[-train_index, ]

#train_data$y_factor <- relevel(train_data$y_factor, ref = "1")
#test_data$y_factor  <- relevel(test_data$y_factor, ref = "1")

```

## Logistic Regression

```{R}
model_log <- glm(y_factor ~ ., 
                 data = train_data,
                 family = binomial)

prob_log <- predict(model_log, test_data, type = "response")

# Reverse threshold
pred_log <- ifelse(prob_log > 0.5, 0, 1)

pred_log <- factor(pred_log, levels = c(0,1))

confusionMatrix(pred_log, test_data$y_factor)
```

## Decision Tree

```{R}
model_tree <- rpart(y_factor ~ ., 
                    data = train_data,
                    method = "class")

rpart.plot(model_tree)

pred_tree <- predict(model_tree, test_data, type = "class")

confusionMatrix(pred_tree, test_data$y_factor)
```

## Random Forest

```{R}
# I'm not sure where this came from, but it seems to not work
#varImpPlot(model_rf) # check for near zero importance variables if yes then delete them and retrain model

model_rf <- randomForest(y_factor ~ ., 
                         data = train_data,
                         ntree = 200)
pred_rf <- predict(model_rf, test_data)

confusionMatrix(pred_rf, test_data$y_factor)

rf_probs <- predict(model_rf, test_data, type = "prob")[,2]
roc_rf <- roc(test_data$y_factor, rf_probs)

auc(roc_rf)
plot(roc_rf)
```

# Model Evaluation

## Calculate Logistic Regression probabilities (probability of class 1)

```{R}
log_probs <- predict(model_log, test_data, type = "response")
roc_log  <- roc(test_data$y_factor, log_probs)
```

## Decision Tree probabilities

```{R}
tree_probs <- predict(model_tree, test_data, type = "prob")[,2]
roc_tree <- roc(test_data$y_factor, tree_probs)

```

## Random Forest probabilities

```{R}
rf_probs <- predict(model_rf, test_data, type = "prob")[,2]
roc_rf   <- roc(test_data$y_factor, rf_probs)
```


## Plot All ROC Curves on One Graph

```{R}
plot(roc_log, 
     col = "blue", 
     lwd = 2,
     main = "ROC Curve Comparison")

plot(roc_tree, 
     col = "green", 
     lwd = 2,
     add = TRUE)

plot(roc_rf, 
     col = "red", 
     lwd = 2,
     add = TRUE)

legend("bottomright",
       legend = c(
         paste("Logistic Regression (AUC =", round(auc(roc_log),3),")"),
         paste("Decision Tree (AUC =", round(auc(roc_tree),3),")"),
         paste("Random Forest (AUC =", round(auc(roc_rf),3),")")
       ),
       col = c("blue", "green", "red"),
       lwd = 2)
```


# Marketing Campaign Strategy

Based on the modeling results:

-   Random Forest achieved the highest AUC and strongest predictive power.
-   High-frequency coffee visitors show highest responsiveness.
-   Time-of-day effects suggest targeted push timing.
-   Coupon type performance varies significantly.

### Proposed Campaign:

1.  Target high-frequency coffee consumers.
2.  Schedule promotions during high-acceptance time windows.
3.  Allocate more budget to top-performing coupon categories.
4.  Use predictive scores to send personalized offers.

------------------------------------------------------------------------

------------------------------------------------------------------------


```{r}
library(randomForest)
library(caret)

# Create interactions manually
train_data$coffee_coupon <- ifelse(train_data$coupon=="Coffee House", train_data$coffee_house, 0)
train_data$restaurant20_coupon <- ifelse(train_data$coupon=="Restaurant(20-50)", train_data$restaurant20to50, 0)
train_data$restaurant_lt20_coupon <- ifelse(train_data$coupon=="Restaurant(<20)", train_data$restaurant_less_than20, 0)
train_data$bar_coupon <- ifelse(train_data$coupon=="Bar", train_data$bar, 0)

train_data$time_coupon <- interaction(train_data$time, train_data$coupon)
train_data$weather_coupon <- interaction(train_data$weather, train_data$coupon)
train_data$passanger_coupon <- interaction(train_data$passanger, train_data$coupon)

# Create interactions manually
test_data$coffee_coupon <- ifelse(test_data$coupon=="Coffee House", test_data$coffee_house, 0)
test_data$restaurant20_coupon <- ifelse(test_data$coupon=="Restaurant(20-50)", test_data$restaurant20to50, 0)
test_data$restaurant_lt20_coupon <- ifelse(test_data$coupon=="Restaurant(<20)", test_data$restaurant_less_than20, 0)
test_data$bar_coupon <- ifelse(test_data$coupon=="Bar", test_data$bar, 0)


test_data$time_coupon <- interaction(test_data$time, test_data$coupon)
test_data$weather_coupon <- interaction(test_data$weather, test_data$coupon)
test_data$passanger_coupon <- interaction(test_data$passanger, test_data$coupon)
```

```{r}
set.seed(123) # for reproducibility

rf_model <- randomForest(
  y ~ destination + passanger + weather + time + coupon +
      expiration + gender + age + marital_status + income +
      coffee_coupon + restaurant_lt20_coupon + restaurant20_coupon + bar_coupon +
      direction_same,
  data = train_data,
  ntree = 1000,
  mtry = floor(sqrt(ncol(train_data)-1)), # typical heuristic
  importance = TRUE
)
```

```{R}
# Probabilities
p_rf <- predict(rf_model, test_data, type="prob")[,2]

# Tune threshold if needed
threshold <- 0.5
pred_rf <- factor(ifelse(p_rf > threshold, 1, 0), levels = levels(test_data$y))

# Confusion matrix
confusionMatrix(pred_rf, test_data$y)
```
GLM "Improved"
```{R}
model_glm_improved <- glm(
  y ~ destination + passanger + weather + time + coupon +
      expiration + gender + marital_status + education + occupation + income +
      coffee_coupon + restaurant_lt20_coupon + restaurant20_coupon + bar_coupon +
      time_coupon + weather_coupon + passanger_coupon +
      direction_same,
  data = train_data,
  family = binomial
)
```
```{R}
p <- predict(model_glm_maxed, test_data, type="response")

# Threshold tuning
threshold <- 0.48  # tweak slightly for best balanced accuracy
pred <- factor(ifelse(p > threshold, 1, 0), levels = levels(test_data$y))

library(caret)
confusionMatrix(pred, test_data$y)
```