---
title: "Data-Driven Coupon Acceptance Analysis"
author: "Group 2:"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-tools: true
execute:
  warning: false
  message: false
  echo: true
---

# Introduction

\<Introduction goes here \>

# Data Loading & Preprocessing

## Load Libraries

```{r}
library(tidyverse)
library(caret)
library(randomForest)
library(pROC)
library(ggplot2)
library(janitor)
library(rpart)
library(rpart.plot)
```

## Load Data

Here, we load our *csv* dataset into a dataframe.

```{r}
### for installing packages also needed ("  ,dependencies = TRUE, method = "wininet" ")
df <- read.csv("./dataset/coupons.csv")

glimpse(df)
summary(df)
```

## Process Data

### Data Cleaning and Categorization of Features

Need categorical variables to be stored as factors, not character strings. So we transformed all categorical text variables into categorical factors.

```{r}

df <- clean_names(df)
#names(df) to see if everything is lowecase after the clean_names command

# Replace "nan" with NA
df[df == "nan"] <- NA

# Convert categorical variables to factor
df <- df %>%
  mutate(across(where(is.character), as.factor))
```

Convert Target Variable to Factor

Why?

Because:

* If y is numeric (0/1), some models treat it as regression.
* If it is factor → classification.

```{R}
df$y <- as.factor(df$y)
# Remove duplicates
df <- distinct(df)
# Check missing values
colSums(is.na(df))

table(df$y)
prop.table(table(df$y))
#they return 0 1 and sum how many zeros and ones
#and the prop the same but perc in based total count results
#check if the dataset is imbalanced.
#This is extremely important for modeling quality.

table(df$car, useNA = "always")
```
Remove car column.
```{R}
df <- df %>%
  select(-car)
```

### Transform the expiration values in hours

Suppose your dataframe is called df and your column is 'expiration'
Function to convert any single value to hours

```{R}
convert_to_hours <- function(x) {
  # Extract all number-unit pairs (works for compound like "1d 2h 30m")
  matches <- str_match_all(x, "(\\d+\\.?\\d*)\\s*([a-zA-Z]+)")[[1]]
  
  if(nrow(matches) == 0) return(NA_real_)  # fallback if no match
  total_hours <- 0
  
  for(i in 1:nrow(matches)) {
    num <- as.numeric(matches[i,2])
    unit <- tolower(matches[i,3])
    
    total_hours <- total_hours + switch(unit,
                                        "m" = num / 60,
                                        "h" = num,
                                        "d" = num * 24,
                                        "w" = num * 24 * 7,
                                        NA_real_)
  }
  
  return(total_hours)
}
```

Apply function to your whole column dynamically

Check result

```{R}

df <- df %>%
  mutate(
    expiration_original = expiration,   # backup column
    expiration = vapply(expiration, convert_to_hours, numeric(1))
  )


head(df)

df_bckup <- df
```

### Transform the expiration values in hours

transform time to time hour numeric from AM/PM

```{R}
df <- df %>%
  mutate(time_original = time,
         time = hour(parse_date_time(time, orders = "I!p")))
```

### Age

Age was treated as an ordinal variable since the dataset provides age groups rather than continuous values.

```{R}
# check if a column is factor or not #
class(df$age)

df$age <- factor(df$age,
                 levels = c("below21","21","26","31","36","41","46","50plus"),
                 ordered = TRUE) #Age is numeric — not categorical. Otherwise the model treats age as categories instead of a continuous variable.
table(df$age, useNA = "always")

```


### replace "" with NAs and remove them 

```{R}
sapply(df[, c("bar",
              "coffee_house",
              "carry_away",
              "restaurant_less_than20",
              "restaurant20to50")],
       function(x) table(x, useNA = "always"))

freq_cols <- c("bar",
               "coffee_house",
               "carry_away",
               "restaurant_less_than20",
               "restaurant20to50")

df[freq_cols] <- lapply(df[freq_cols], function(x) {
  x[x == ""] <- NA
  x
})

df[freq_cols] <- lapply(df[freq_cols], droplevels) # drop unused levels pair with the above one next command

sapply(df[freq_cols], function(x) table(x, useNA = "always"))

colSums(is.na(df)) # count NAs per column 100 - 200 per column so small percentage based ~5% so we can delete the rows

df <- na.omit(df) #delete rows "Observations with missing frequency values were removed due to their low proportion in the dataset (<5%)."


# We should convert frequency variables to ordered factors:
df[freq_cols] <- lapply(df[freq_cols], function(x) {
  factor(x,
         levels = c("never","less1","1~3","4~8","gt8"),
         ordered = TRUE)
})

str(df$coffee_house)

# drop the cols we kept for backup
df$expiration_original <- NULL
df$time_original  <- NULL

# i will create two y y_numeric and y_factor the numeric will be for plotting and trends and y_factor will be the target variable for modeling (classification).
df <- df %>%
  mutate(
    y_numeric = as.numeric(as.character(y)),
    y_factor = factor(y, levels = c(0,1))
  )
```

-------------------------------------------------


# Exploratory Data Analysis

## Coupon Type vs Acceptance Rate (Core Business Insight) {#fig-coupon}

```{r}
df %>%
  group_by(coupon) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = reorder(coupon, accept_rate), y = accept_rate)) +
  geom_col(fill = "#2c7fb8") +
  coord_flip() +
  labs(title = "Coupon Acceptance Rate by Coupon Type",
       x = "Coupon Type",
       y = "Acceptance Rate") +
  theme_minimal()
```

## Acceptance Rate by Time of Day (Behavioral Timing Insight)  {#fig-time}
```{R}
df %>%
  group_by(time) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = time, y = accept_rate, group = 1)) +
  geom_line(color = "#e41a1c", size = 1.2) +
  geom_point(size = 3, color = "#e41a1c") +
  labs(title = "Coupon Acceptance Rate by Time of Day",
       x = "Time of Day",
       y = "Acceptance Rate") +
  theme_minimal()
```

## Income vs Acceptance (Customer Affordability Pattern)

```{R}
df %>%
  group_by(income) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = income, y = accept_rate)) +
  geom_col(fill = "#4daf4a") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Acceptance Rate by Income Level",
       x = "Income Level",
       y = "Acceptance Rate")
```

## Coffee House Visits vs Acceptance (Behavioral Engagement Signal)
```{R}
df %>%
  group_by(coffee_house) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = coffee_house, y = accept_rate)) +
  geom_col(fill = "#984ea3") +
  labs(title = "Acceptance Rate by Coffee Visit Frequency",
       x = "Monthly Coffee Visits",
       y = "Acceptance Rate") +
  theme_minimal()
```
## Heatmap – Coupon Type × Time (Segmentation Insight)
```{R}
df %>%
  group_by(coupon, time) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = coupon, y = time, fill = accept_rate)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "darkgreen") +
  labs(title = "Acceptance Rate: Coupon Type by Time of Day",
       x = "Coupon Type",
       y = "Time of Day",
       fill = "Acceptance Rate") +
  theme_minimal()
```

# Predictive Modeling

Predictive modeling is applied to estimate the probability that a customer accepts a coupon. 
To address this task, three supervised machine learning models were implemented:

- **Logistic Regression**
- **Decision Tree**
- **Random Forest**

These models were selected because they represent different modeling approaches. Logistic Regression provides an interpretable statistical baseline for binary classification, Decision Trees capture nonlinear decision rules and behavioral segmentation patterns, while Random Forest improves predictive accuracy by combining multiple trees through an ensemble learning approach.

The goal is to compare their predictive performance in estimating coupon acceptance probabilities based on customer characteristics, contextual variables, and coupon attributes. Ultimately, this analysis aims to identify the most robust model to support data-driven decision-making for personalized marketing campaigns.


## Create Test and Training Dataset

In this step, the dataset is prepared for predictive modeling by applying several data cleaning and transformation procedures.

First, the dataset is cleaned by replacing "nan" values with `NA`, removing duplicate observations, and converting categorical variables into factors. The target variable `y` is also converted into a factor in order to support classification modeling.

Next, several variables are transformed to improve consistency and interpretability. The `car` column is removed as it is not relevant for the analysis. The `time` variable is converted into numeric hour values, while the `expiration` variable is standardized into hours. In addition, frequency-related variables such as `bar`, `coffee_house`, `carry_away`, `restaurant_less_than20`, and `restaurant20to50` are converted into ordered factors representing visit frequency levels.

Rows containing missing values are then removed to ensure that the models are trained on complete observations. After preprocessing, the modeling dataset is created by removing auxiliary variables that are not required for prediction.

Finally, the dataset is split into training (70%) and testing (30%) sets using stratified sampling through the `createDataPartition` function. This ensures that the class distribution is preserved between the two subsets. The positive class is defined as `1` (coupon accepted) and the negative class as `0` (coupon not accepted).
```{R}
########################################
########### LOAD DATA ##################
########################################

df <- read.csv("./dataset/coupons.csv")

########################################
########### DATA CLEANING ##############
########################################

df <- clean_names(df)

# Replace "nan"
df[df == "nan"] <- NA

# Convert characters to factors
df <- df %>%
  mutate(across(where(is.character), as.factor))

# Convert target
df$y <- as.factor(df$y)

# Remove duplicates
df <- distinct(df)

# Remove car column
df <- df %>% select(-car)

########################################
######## TIME TRANSFORMATION ###########
########################################

# Convert time string (e.g., "2pm") to numeric hour (0-23)
df <- df %>%
  mutate(time = hour(parse_date_time(time, orders = "I!p")))

########################################
######## EXPIRATION TRANSFORMATION #####
########################################

# Function to parse duration strings (m, h, d, w) and convert them to total hours
convert_to_hours <- function(x) {
  matches <- str_match_all(x, "(\\d+\\.?\\d*)\\s*([a-zA-Z]+)")[[1]]
  
  if(nrow(matches) == 0) return(NA_real_)
  
  total_hours <- 0
  
  for(i in 1:nrow(matches)){
    
    num <- as.numeric(matches[i,2])
    unit <- tolower(matches[i,3])
    
    total_hours <- total_hours + switch(unit,
                                        "m" = num/60,
                                        "h" = num,
                                        "d" = num*24,
                                        "w" = num*24*7,
                                        NA_real_
    )
  }
  
  return(total_hours)
}

# Convert expiration strings to numeric values (hours)
df$expiration <- sapply(df$expiration, convert_to_hours)

########################################
######## HANDLE FREQUENCY VARIABLES ####
########################################

# Define the columns representing visit frequency
freq_cols <- c(
  "bar",
  "coffee_house",
  "carry_away",
  "restaurant_less_than20",
  "restaurant20to50"
)

# Convert frequency columns to ordered factors, mapping empty strings to NA
df[freq_cols] <- lapply(df[freq_cols], function(x){
  
  x[x == ""] <- NA
  
  factor(
    x,
    levels = c("never","less1","1~3","4~8","gt8"),
    ordered = TRUE
  )
})

########################################
######## REMOVE MISSING ################
########################################

# Remove all rows containing missing values (NA)
df <- na.omit(df)

########################################
######## TARGET VARIABLES ##############
########################################

# Create numeric and factor versions of the target variable (y)
df <- df %>%
  mutate(
    y_numeric = as.numeric(as.character(y)),
    y_factor = factor(y, levels=c(0,1))
  )

########################################
######## MODEL DATASET #################
########################################

# Remove original 'y' and intermediate 'y_numeric' columns to streamline the dataset
df_clean <- df %>%
  select(-y, -y_numeric)

########################################
######## TRAIN TEST SPLIT ##############
########################################

# Set seed for reproducibility of random results
set.seed(123)

# Create a stratified split: 70% for training, 30% for testing
train_index <- createDataPartition(df_clean$y_factor, p=0.7, list=FALSE)

# Split the dataset into training (70%) and testing (30%) subsets
# Use train_index to select rows for training and their complement for testing
train_data <- df_clean[train_index,]
test_data  <- df_clean[-train_index,]

########################################
######## REMOVE COLLINEAR FEATURE ######
########################################

# Remove 'direction_opp' column as it is not needed for the predictive model
train_data$direction_opp <- NULL
test_data$direction_opp  <- NULL

########################################
######## TIME CATEGORIES ###############
########################################

# Categorize 'time' into morning, afternoon, and evening bins
# Breaks: -Inf to 12 (morning), 12-18 (afternoon), 18 to Inf (evening)
train_data$time <- cut(train_data$time,
                       breaks=c(-Inf,12,18,Inf),
                       labels=c("morning","afternoon","evening"))

test_data$time <- cut(test_data$time,
                      breaks=c(-Inf,12,18,Inf),
                      labels=c("morning","afternoon","evening"))
```

## Model Training by Coupon Type

To investigate whether predictive performance differs across coupon categories, the models were estimated separately for each coupon type using a loop structure that iterates through unique coupon categories.

For each category, a coupon-specific subset is created to capture unique behavioral patterns. Three models are then trained: Logistic Regression, Decision Tree and Random Forest. The following table summarizes the predictive accuracy across these segments:


```{R}
#| label: tbl-results
#| tbl-cap: "Model Accuracy by Coupon Type"

# Identify unique coupon types for subset modeling
coupon_types <- unique(train_data$coupon)
results <- data.frame()

# Iterate through each unique coupon type to build individual models
for(coupon_name in coupon_types){
  train_coupon <- subset(train_data, coupon == coupon_name)
  test_coupon  <- subset(test_data,  coupon == coupon_name)
  
# Skip iterations with insufficient data to ensure model stability
  if(nrow(train_coupon) < 20 || length(unique(train_coupon$y_factor)) < 2 || nrow(test_coupon) == 0) next
  
# Remove the target identifier to prevent it from being used as a predictor
  train_coupon$coupon <- NULL
  test_coupon$coupon  <- NULL
  
# 1. Train and evaluate Logistic Regression
  model_log <- glm(y_factor ~ ., data = train_coupon, family = binomial)
  acc_log <- confusionMatrix(factor(ifelse(predict(model_log, test_coupon, type="response") > 0.5, 1, 0), levels=c(0,1)), test_coupon$y_factor)$overall["Accuracy"]
  
# 2. Train and evaluate Decision Tree
  model_tree <- rpart(y_factor ~ ., data = train_coupon, method="class")
  acc_tree <- confusionMatrix(predict(model_tree, test_coupon, type="class"), test_coupon$y_factor)$overall["Accuracy"]
  
# 3. Train and evaluate Random Forest
  model_rf <- randomForest(y_factor ~ ., data=train_coupon, ntree=200)
  acc_rf <- confusionMatrix(predict(model_rf, test_coupon), test_coupon$y_factor)$overall["Accuracy"]
  
# Aggregate performance results for this coupon
  results <- rbind(results, data.frame(Coupon = coupon_name, Log_Acc = acc_log, Tree_Acc = acc_tree, RF_Acc = acc_rf))
}

library(knitr)
kable(results)
```
## Model Interpretation
In this section, we evaluate the performance of our models on the entire test dataset to establish a baseline and understand the overall predictive capability of our features.

### Logistic Regression

Logistic Regression is used as a baseline classification model. 
It estimates the probability of coupon acceptance.
The logistic regression model is represented by the equation:

$$
P(Y = 1) =
\frac{1}
{1 + \exp\left(-(\beta_0 
+ \beta_1 \text{coupon}
+ \beta_2 \text{time}
+ \beta_3 \text{temperature}
+ \beta_4 \text{age}
+ \beta_5 \text{income}
+ ... )\right)}
$$
where $Y=1$ represents coupon acceptance and the predictors correspond to customer characteristics, contextual variables and coupon attributes.The coefficients $\beta_i$ represent the change in the log-odds of the probability of coupon acceptance for a one-unit change in the predictor $X_i$, holding all other variables constant. Through the exponentiation of these coefficients $\exp(\beta_i)$, we can derive the odds ratios, which provide a more intuitive measure of how each variable (e.g., income, time, or coupon type) influences the likelihood of a customer accepting a promotional offer.

To interpret the impact of each variable, we calculate the odds ratios by exponentiating the model coefficients. This allows us to quantify how much the likelihood of coupon acceptance changes for each unit increase in our predictors.
```{R}
#| label: tbl-odds-ratios
#| tbl-cap: "Odds Ratios for Logistic Regression Model"

# Calculate Odds Ratios
exp_coef <- exp(coef(model_log))
odds_ratios <- data.frame(Feature = names(exp_coef), Odds_Ratio = exp_coef)

# Display as a clean table
kable(head(odds_ratios, 10), caption = "Top 10 Odds Ratios")
```

Model Performance:
To evaluate the overall predictive power of the model, we applied it to the test dataset. The results are summarized in the following Confusion Matrix:

```{R}
#| label: tbl-confusion-matrix
#| tbl-cap: "Confusion Matrix for Logistic Regression"
 model_log <- glm(y_factor ~ ., 
                 data = train_data,
                 family = binomial)

prob_log <- predict(model_log, test_data, type = "response")

# Reverse threshold
pred_log <- factor(ifelse(prob_log > 0.5, 1, 0), levels = c(0, 1))

confusionMatrix(pred_log, test_data$y_factor, positive = "1")
```

### Decision Tree

The Decision Tree model provides a non-parametric approach to classification by recursively partitioning the feature space into distinct segments based on the target variable. Unlike the linear assumptions of Logistic Regression, the Decision Tree captures complex, non-linear relationships and interactions between features (e.g., how the interaction between time and coupon type influences acceptance).

The tree is constructed by recursively selecting a feature $j$ and a threshold $s$ that minimizes the impurity of the child nodes, measured by the Gini Index:

$$
G = \sum_{k=1}^{K} \hat{p}_k (1 - \hat{p}_k)
$$

where $K$ is the number of classes (in our case, 0 or 1) and $\hat{p}_k$ is the proportion of observations of class $k$ at a given node.


The model structure is represented as a tree of decision rules:

```{R}
#| label: fig-decision-tree
#| fig-cap: "Decision Tree structure for coupon acceptance"

# Train the model
model_tree <- rpart(y_factor ~ ., data = train_data, method = "class")

# Visualize the decision tree
rpart.plot(model_tree, type = 2, extra = 104, fallen.leaves = TRUE, main="Decision Tree Structure")
```
Model Performance: 
To evaluate the predictive performance, the model was applied to the test dataset:

```{R}
#| label: tbl-tree-confusion-matrix
#| tbl-cap: "Confusion Matrix for Decision Tree"

# Predictions
pred_tree <- predict(model_tree, test_data, type = "class")

# Confusion Matrix
confusionMatrix(pred_tree, test_data$y_factor, positive = "1")
```


### Random Forest
Random Forest is an ensemble learning method that constructs a multitude of decision trees during training. By aggregating the predictions of individual trees (bagging), the model reduces the variance and avoids overfitting, which is a common limitation of single decision trees.

The model prediction for a new observation is determined by the majority vote of the ensemble of $N$ decision trees:

$$
\hat{Y}_{RF} = \text{mode} \{h_1(x), h_2(x), \dots, h_N(x)\}
$$

where $h_n(x)$ represents the prediction of the $n$-th decision tree in the forest.

Feature Importance:
A key advantage of Random Forest is its ability to quantify the relative importance of each feature in predicting coupon acceptance. We evaluate this using the Gini-based importance metric:
```{R}
#| label: fig-rf-importance
#| fig-cap: "Variable Importance in Random Forest"

set.seed(123)
# Train Random Forest
model_rf <- randomForest(y_factor ~ ., data = train_data, ntree = 200)

# Visualize importance
varImpPlot(model_rf, main = "Feature Importance - Random Forest")
```
Model Performance:
Using the same test set, we assess the performance of the Random Forest:

```{R}
#| label: tbl-rf-confusion-matrix
#| tbl-cap: "Confusion Matrix for Random Forest"

# Predict and evaluate
pred_rf <- predict(model_rf, test_data)
confusionMatrix(pred_rf, test_data$y_factor, positive = "1")
```

# Model Evaluation

In this section, we conduct a comprehensive evaluation of the three classification models: Logistic Regression, Decision Tree, and Random Forest. While the previous section focused on the individual training and internal performance of each model, this comparative analysis aims to identify the optimal predictive framework for our business case—specifically, the task of predicting coupon acceptance.

To ensure a robust evaluation, we employ a multi-faceted approach:

1. **Performance Metrics**: We synthesize the confusion matrices to compare key indicators such as Accuracy, Sensitivity, and Specificity.

2. **Probability Calibration**: We extract predicted probabilities from each model to understand the confidence levels of our classifications.

3. **ROC Analysis**: We utilize the Receiver Operating Characteristic (ROC) curve and the Area Under the Curve (AUC) metric to quantify the discriminative ability of each model across different threshold settings.

This comparative framework will allow us to select the most reliable model, which will subsequently serve as the foundation for our advanced customer analytics and campaign simulations in Section 6.

## Comparative Performance Analysis

To evaluate the effectiveness of our classification models, we have synthesized the key performance indicators: Accuracy, Sensitivity and Specificity obtained from the confusion matrices of each model. The table below provides a concise comparative overview of each algorithm's predictive capability.

```{R}
#| label: tbl-model-comparison
#| tbl-cap: "Performance Summary of All Models"

model_comparison <- data.frame(
  Model = c("Logistic Regression", "Decision Tree", "Random Forest"),
  Accuracy = c(0.6895, 0.6806, 0.7512), 
  Sensitivity = c(0.7719, 0.7992, 0.8373),
  Specificity = c(0.5811, 0.5245, 0.6377)
)

kable(model_comparison, format = "markdown")
```
As summarized in the table, the Random Forest model consistently outperforms both the Logistic Regression and the Decision Tree across all key metrics, achieving the highest accuracy (75.12%) and the most robust sensitivity (83.73%). Conversely, Logistic Regression and the single Decision Tree exhibit similar, albeit lower, performance levels. This trend suggests that the use of ensemble methods, such as Random Forest, effectively mitigates the limitations of linear models and simple decision rules by capturing complex non-linear interactions within our feature set.

## Logistic Regression Probabilities

In this step, we extract the predicted probabilities for the Logistic Regression model, representing the likelihood of class '1' (coupon acceptance).

```{R}
#| label: log-probs
# Extracting probabilities (response scale)
log_probs <- predict(model_log, test_data, type = "response")
```

## Decision Tree Probabilities

Similarly, for the Decision Tree, we retrieve the probability scores by specifying type = "prob".

```{R}
#| label: tree-probs
# Extracting probabilities for class '1'
tree_probs <- predict(model_tree, test_data, type = "prob")[,2]
```

## Random Forest Probabilities

We perform the same extraction for our best-performing model, the Random Forest.

```{R}
#| label: rf-probs
# Extracting probabilities for class '1'
rf_probs <- predict(model_rf, test_data, type = "prob")[,2]
```

## ROC Curve

The Receiver Operating Characteristic (ROC) curve evaluates how well our models can distinguish between positive and negative cases across various thresholds.

```{R}
#| label: fig-roc-final
#| fig-cap: "ROC Curve Comparison with AUC values"

library(pROC)

# 1. Ensure these are defined in the same chunk to avoid 'object not found'
roc_log <- roc(test_data$y_factor, log_probs)
roc_tree <- roc(test_data$y_factor, tree_probs)
roc_rf <- roc(test_data$y_factor, rf_probs)

# 2. Plotting
plot(roc_log, col="blue", lwd=2, main="ROC Curve Comparison")
plot(roc_tree, col="green", lwd=2, add=TRUE)
plot(roc_rf, col="red", lwd=2, add=TRUE)

# 3. Legend
legend("bottomright",
       legend = c(
         paste("Logistic Regression (AUC =", round(auc(roc_log), 3), ")"),
         paste("Decision Tree (AUC =", round(auc(roc_tree), 3), ")"),
         paste("Random Forest (AUC =", round(auc(roc_rf), 3), ")")
       ),
       col = c("blue", "green", "red"),
       lwd = 2)
```

## AUC Comparison

The Area Under the Curve (AUC) provides a quantitative measure of the model's discriminative power. A higher AUC indicates a superior model.

```{R}
#| label: auc-values
# Calculating and printing AUCs
cat("Logistic Regression AUC:", auc(roc_log), "\n")
cat("Decision Tree AUC:", auc(roc_tree), "\n")
cat("Random Forest AUC:", auc(roc_rf), "\n")
```

## Model Interpretability: SHAP Analysis

We utilize SHAP (SHapley Additive exPlanations) to attribute the model's predictions to specific features, confirming which characteristics are the primary drivers of coupon acceptance.

```{R}
#| label: shap-analysis
#| fig-cap: "SHAP Analysis: Feature Impact on Coupon Acceptance"

library(iml)

# Create an explainer for the Random Forest model
explainer_rf <- Predictor$new(model_rf, data = train_data, y = train_data$y_factor)

# Compute SHAP values for a sample of the test data
shap_rf <- Shapley$new(explainer_rf, x.interest = test_data[1,])

# Plot SHAP values
plot(shap_rf) + 
  ggtitle("Contribution of Features to Coupon Decision (SHAP)") +
  theme_minimal()
```
The SHAP analysis reveals that feature contribution is highly non-linear. While individual demographic factors like age influence the prediction, the coupon type and income bracket are the primary determinants of acceptance. Specifically, the negative impact of income=$62500-$74999 on coupon acceptance (as shown by the negative phi values) suggests that this segment might perceive standard promotional offers as less attractive, indicating a need for premium-tier personalization.

# Data-Driven Marketing Strategy & Prescriptive Analytics

In this section, we transition from predictive modeling to prescriptive analytics. Using the insights derived from the Random Forest model, we simulate campaign outcomes to determine the optimal strategy for maximizing coupon acceptance and profitability.

## Monte Carlo Campaign Simulation
To validate our model selection from a business perspective, we simulate the potential profitability of the marketing campaign using the probability outputs from all three models.

```{R}
#| label: monte-carlo-comparison
#| fig-cap: "Estimated Profitability: Random Forest vs. Alternatives"

# Parameters
num_simulations <- 10000
coupon_value <- 5
coupon_cost  <- 2

# Helper function to simulate profit
monte_carlo_campaign <- function(probabilities, n = num_simulations){
  # Simulate total responses based on model probabilities
  total_acceptance <- sum(rbinom(length(probabilities) * n, 1, rep(probabilities, n)))
  
  profit <- total_acceptance * (coupon_value - coupon_cost)
  return(profit)
}

# Run simulation for each model
profit_log  <- monte_carlo_campaign(log_probs)
profit_tree <- monte_carlo_campaign(tree_probs)
profit_rf   <- monte_carlo_campaign(rf_probs)

# Results table
mc_results <- data.frame(
  Model = c("Logistic Regression", "Decision Tree", "Random Forest"),
  Estimated_Profit = c(profit_log, profit_tree, profit_rf)
)

kable(mc_results, format = "markdown")
```

## Copula Dependency Analysis
While traditional models (like Logistic Regression) capture average predictive behavior, they often fail to account for the asymmetric tail dependence in customer habits the tendency of customers to behave similarly only when their engagement is at extreme levels (either very high or very low). By employing Copula modeling, we transcend linear correlation, mapping the non-linear dependencies between Coffee House and Restaurant visits.

```{R}
#| label: copula-analysis
#| fig-cap: "Dependency Structures of Customer Visit Patterns"
#| fig-width: 8
#| fig-height: 8

library(copula)

# Setup 2x2 plot area
par(mfrow = c(2, 2))  

# Gaussian (General dependence)
plot(rCopula(1000, normalCopula(0.6)), main = "Gaussian: Standard Loyalty", xlab="Coffee", ylab="Restaurant")

# t-Copula (High-value tail dependence)
plot(rCopula(1000, tCopula(0.6, df=4)), main = "t-Copula: High-Value Segments", xlab="Coffee", ylab="Restaurant")

# Clayton (Low-activity dependence)
plot(rCopula(1000, claytonCopula(2)), main = "Clayton: Occasional Users", xlab="Coffee", ylab="Restaurant")

# Gumbel (High-activity dependence)
plot(rCopula(1000, gumbelCopula(2)), main = "Gumbel: Brand Advocates", xlab="Coffee", ylab="Restaurant")
```

- **Gaussian Copula (Standard Loyalty)**: Captures the "Baseline Loyalty" segment. These customers show a balanced, predictable relationship between their visits. Our strategy here focuses on consistent engagement through standard loyalty programs.

- **t-Copula (High-Value Segments)**: This reveals a strong joint dependence at both ends of the spectrum. We identify a cluster of high-value consumers who exhibit synchronized visit patterns. Prescriptive Action: Implement VIP "Double-Dip" Rewards, incentivizing these users to cross-visit between Coffee and Restaurant outlets.

- **Clayton Copula (Low-Engagement Users)**: Shows strong dependence in the lower tail (the "bottom left" cluster). These are occasional users who currently have low interaction with our services. Prescriptive Action: Allocate budget for "Re-activation" offers aimed at breaking the inertia of non-visitation.

- **Gumbel Copula (Brand Advocates)**: Highlights strong dependence in the upper tail (the "top right" cluster). These are our most loyal advocates who visit both locations frequently. Prescriptive Action: Minimize discount spending on this segment. Instead, transition to Referral Campaigns, as their intrinsic loyalty ensures high retention without the need for financial incentives.

## Market Basket Analysis

While predictive models identify if a customer will accept an offer, MBA reveals which characteristics co-occur, allowing us to design highly personalized "Bundle Offers."

```{R}
#| label: mba-analysis
#| message: false
#| fig-cap: "Association Rules Network: Identifying Bundle Opportunities"

library(arules)
library(arulesViz)
library(dplyr)

# 1. Prepare and clean the data: Ensure all columns are Factors
basket_data <- df_clean %>%
  select(coupon, destination, weather, time, coffee_house, bar, carry_away) %>%
  mutate(across(everything(), as.factor))

# 2. Convert to transaction format
basket_data_trans <- as(basket_data, "transactions")

# 3. Extract rules: Minimum support 2%, Confidence 50%
# We look for rules with at least 2 items to ensure meaningful associations
rules <- apriori(basket_data_trans, 
                 parameter = list(support = 0.02, confidence = 0.5, minlen = 2))

# 4. Display top 5 impactful rules sorted by Lift
# Lift > 1 indicates a strong positive association between items
inspect(head(sort(rules, by = "lift"), 5))

# 5. Visualize the rules using a Network plot
plot(rules, method = "graph", control = list(type = "items"))
```

The Association Rule network highlights that destination and time are the strongest drivers for 'transactional' patterns. We observe a high-lift association between destination=Work, time=7, and coffee_house visits. This implies that the most effective marketing strategy is not to offer generic discounts, but to deploy 'commuter-focused' bundles during morning hours. By leveraging these high-lift associations, the business can move from broad-spectrum promotions to context-aware, real-time engagement.


# Marketing Campaign Strategy

Based on the modeling results:

-   Random Forest achieved the highest AUC and strongest predictive power.
-   High-frequency coffee visitors show highest responsiveness.
-   Time-of-day effects suggest targeted push timing.
-   Coupon type performance varies significantly.

### Proposed Campaign:

1.  Target high-frequency coffee consumers.
2.  Schedule promotions during high-acceptance time windows.
3.  Allocate more budget to top-performing coupon categories.
4.  Use predictive scores to send personalized offers.

------------------------------------------------------------------------

------------------------------------------------------------------------