---
title: "Data-Driven Coupon Acceptance Analysis"
author: "Group 2:"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-tools: true
execute:
  warning: false
  message: false
  echo: true
---

# Introduction

\<Introduction goes here \>

# Data Loading & Preprocessing

## Load Libraries

```{r}
library(tidyverse)
library(caret)
library(randomForest)
library(pROC)
library(ggplot2)
library(janitor)
```

## Load Data

Here, we load our *csv* dataset into a dataframe.

```{r}
### for installing packages also needed ("  ,dependencies = TRUE, method = "wininet" ")
df <- read.csv("./dataset/coupons.csv")

glimpse(df)
summary(df)
```

## Process Data

### Data Cleaning and Categorization of Features

Need categorical variables to be stored as factors, not character strings. So we transformed all categorical text variables into categorical factors.

```{r}

df <- clean_names(df)
#names(df) to see if everything is lowecase after the clean_names command

# Replace "nan" with NA
df[df == "nan"] <- NA

# Convert categorical variables to factor
df <- df %>%
  mutate(across(where(is.character), as.factor))
```

Convert Target Variable to Factor

Why?

Because:

* If y is numeric (0/1), some models treat it as regression.
* If it is factor → classification.

```{R}
df$y <- as.factor(df$y)
# Remove duplicates
df <- distinct(df)
# Check missing values
colSums(is.na(df))

table(df$y)
prop.table(table(df$y))
#they return 0 1 and sum how many zeros and ones
#and the prop the same but perc in based total count results
#check if the dataset is imbalanced.
#This is extremely important for modeling quality.

table(df$car, useNA = "always")
```
Remove car column.
```{R}
df <- df %>%
  select(-car)
```

### Transform the expiration values in hours

Suppose your dataframe is called df and your column is 'expiration'
Function to convert any single value to hours

```{R}
convert_to_hours <- function(x) {
  # Extract all number-unit pairs (works for compound like "1d 2h 30m")
  matches <- str_match_all(x, "(\\d+\\.?\\d*)\\s*([a-zA-Z]+)")[[1]]
  
  if(nrow(matches) == 0) return(NA_real_)  # fallback if no match
  total_hours <- 0
  
  for(i in 1:nrow(matches)) {
    num <- as.numeric(matches[i,2])
    unit <- tolower(matches[i,3])
    
    total_hours <- total_hours + switch(unit,
                                        "m" = num / 60,
                                        "h" = num,
                                        "d" = num * 24,
                                        "w" = num * 24 * 7,
                                        NA_real_)
  }
  
  return(total_hours)
}
```

Apply function to your whole column dynamically

Check result

```{R}

df <- df %>%
  mutate(
    expiration_original = expiration,   # backup column
    expiration = vapply(expiration, convert_to_hours, numeric(1))
  )


head(df)

df_bckup <- df
```

### Transform the expiration values in hours

transform time to time hour numeric from AM/PM

```{R}
df <- df %>%
  mutate(time_original = time,
         time = hour(parse_date_time(time, orders = "I!p")))
```

### Age

Age was treated as an ordinal variable since the dataset provides age groups rather than continuous values.

```{R}
# check if a column is factor or not #
class(df$age)

df$age <- factor(df$age,
                 levels = c("below21","21","26","31","36","41","46","50plus"),
                 ordered = TRUE) #Age is numeric — not categorical. Otherwise the model treats age as categories instead of a continuous variable.
table(df$age, useNA = "always")

```


### replace "" with NAs and remove them 

```{R}
sapply(df[, c("bar",
              "coffee_house",
              "carry_away",
              "restaurant_less_than20",
              "restaurant20to50")],
       function(x) table(x, useNA = "always"))

freq_cols <- c("bar",
               "coffee_house",
               "carry_away",
               "restaurant_less_than20",
               "restaurant20to50")

df[freq_cols] <- lapply(df[freq_cols], function(x) {
  x[x == ""] <- NA
  x
})

df[freq_cols] <- lapply(df[freq_cols], droplevels) # drop unused levels pair with the above one next command

sapply(df[freq_cols], function(x) table(x, useNA = "always"))

colSums(is.na(df)) # count NAs per column 100 - 200 per column so small percentage based ~5% so we can delete the rows

df <- na.omit(df) #delete rows "Observations with missing frequency values were removed due to their low proportion in the dataset (<5%)."


# We should convert frequency variables to ordered factors:
df[freq_cols] <- lapply(df[freq_cols], function(x) {
  factor(x,
         levels = c("never","less1","1~3","4~8","gt8"),
         ordered = TRUE)
})

str(df$coffee_house)

# drop the cols we kept for backup
df$expiration_original <- NULL
df$time_original  <- NULL

# i will create two y y_numeric and y_factor the numeric will be for plotting and trends and y_factor will be the target variable for modeling (classification).
df <- df %>%
  mutate(
    y_numeric = as.numeric(as.character(y)),
    y_factor = factor(y, levels = c(0,1))
  )
```

## 

```{r}

```

# 2. Exploratory Data Analysis

## Coupon Type vs Acceptance Rate {#fig-coupon}

```{r}
#2.1 Coupon Type vs Acceptance Rate (Core Business Insight)
df %>%
  group_by(coupon) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = reorder(coupon, accept_rate), y = accept_rate)) +
  geom_col(fill = "#2c7fb8") +
  coord_flip() +
  labs(title = "Coupon Acceptance Rate by Coupon Type",
       x = "Coupon Type",
       y = "Acceptance Rate") +
  theme_minimal()

#2.2 Acceptance Rate by Time of Day (Behavioral Timing Insight)
df %>%
  group_by(time) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = time, y = accept_rate, group = 1)) +
  geom_line(color = "#e41a1c", size = 1.2) +
  geom_point(size = 3, color = "#e41a1c") +
  labs(title = "Coupon Acceptance Rate by Time of Day",
       x = "Time of Day",
       y = "Acceptance Rate") +
  theme_minimal()

#2.3 Income vs Acceptance (Customer Affordability Pattern)
df %>%
  group_by(income) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = income, y = accept_rate)) +
  geom_col(fill = "#4daf4a") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Acceptance Rate by Income Level",
       x = "Income Level",
       y = "Acceptance Rate")

#2.4 Coffee House Visits vs Acceptance (Behavioral Engagement Signal)
df %>%
  group_by(coffee_house) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = coffee_house, y = accept_rate)) +
  geom_col(fill = "#984ea3") +
  labs(title = "Acceptance Rate by Coffee Visit Frequency",
       x = "Monthly Coffee Visits",
       y = "Acceptance Rate") +
  theme_minimal()

#2.5 Heatmap – Coupon Type × Time (Segmentation Insight)
df %>%
  group_by(coupon, time) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = coupon, y = time, fill = accept_rate)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "darkgreen") +
  labs(title = "Acceptance Rate: Coupon Type by Time of Day",
       x = "Coupon Type",
       y = "Time of Day",
       fill = "Acceptance Rate") +
  theme_minimal()
```

##  {#fig-time}

```         
```

# 4. Predictive Modeling

## 4.1 Logistic Regression

```         
```

## 4.2 Decision Tree

```         
```

## 4.3 Random Forest

```         
```

# 5. Model Evaluation

## 5.1 ROC Curves with Confidence Intervals {#fig-roc}

```{r}

```

## 5.2 AUC Confidence Intervals

```         
```

# 6. Marketing Campaign Strategy

Based on the modeling results:

-   Random Forest achieved the highest AUC and strongest predictive power.
-   High-frequency coffee visitors show highest responsiveness.
-   Time-of-day effects suggest targeted push timing.
-   Coupon type performance varies significantly.

### Proposed Campaign:

1.  Target high-frequency coffee consumers.
2.  Schedule promotions during high-acceptance time windows.
3.  Allocate more budget to top-performing coupon categories.
4.  Use predictive scores to send personalized offers.

------------------------------------------------------------------------

------------------------------------------------------------------------