---
title: "Data-Driven Coupon Acceptance Analysis"
author: "Group 2: 
Ayse Ciftci (fXXX),
Christos Kapotis (fXXX),
Ioannis Koutsogiannis (fXXX),
Lazar Nenovski (fXXX),
Dallandyshe Shabani (fXXX)
"
date: today
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-tools: true
  pdf:
    documentclass: scrartcl
    pdf-engine: xelatex
    mainfont: Source Sans Pro
    monofont: Fira Code
    fontsize: 11pt
    geometry:
      - top=2.2cm
      - bottom=2.2cm
      - left=2.2cm
      - right=2.2cm
    number-sections: true
    colorlinks: true
    fig-width: 6.5
    fig-height: 4
    fig-align: center
 
    # Stylistic options for the LaTeX render
    header-includes:
      - \usepackage{fontspec}
      - \usepackage{microtype}
      - \usepackage{booktabs}
      - \usepackage{siunitx}
      - \usepackage{caption}
      - \setlength{\parindent}{0pt}
      - \setlength{\parskip}{6pt}
      - \setkomafont{section}{\normalfont\Large\bfseries}
      - \setkomafont{subsection}{\normalfont\large\bfseries}
      - \usepackage{xcolor}
      - \definecolor{inlinecode}{HTML}{7d12ba}
      - \definecolor{inlinebg}{HTML}{f8f9fa}
      - \let\oldtexttt\texttt
      - \renewcommand{\texttt}[1]{\colorbox{
            inlinebg}{\oldtexttt{\textcolor{inlinecode}{#1}}}}
      - \definecolor{codebg}{HTML}{F5F5F5}
      - \definecolor{outbg}{HTML}{E8E8E8}
    tbl-colwidths: auto

highlight-style: github

execute:
  warning: false
  message: false
  # echo: true
---

# Introduction {#sec-introduction}


The objective of this report is to analyse customer coupon acceptance behavior 
and predict whether a customer will accept a coupon offer. Since the target 
variable has two possible outcomes, accepted or rejected, the problem is 
formulated as a binary classification task. To address this objective, we first 
perform data cleaning and pre-processing to ensure data quality and proper 
variable formatting. We then conduct exploratory data analysis to identify key 
behavioral and contextual patterns associated with coupon acceptance. Finally, 
we develop and compare three predictive models, Logistic Regression, Decision 
Tree, and Random Forest, to evaluate their predictive performance and derive 
actionable marketing insights. The final goal is to identify the customer and 
situational factors that increase coupon acceptance and translate these findings
into actionable marketing recommendations.

# Data Loading & Preprocessing {#sec-data-loading}

## Load Libraries

Command to load the R libraries we will use for the analysis:

```{r, library-setup}
library(tidyverse)
library(caret)
library(randomForest)
library(pROC)
library(ggplot2)
library(janitor)
library(rpart)
library(rpart.plot)
library(iml)
library(knitr)
library(copula)
library(arules)
library(arulesViz)
```

## Load Data

The dataset was imported from a CSV file containing survey responses related to 
coupon acceptance behaviour. Each observation represents a customer scenario 
including demographic attributes, contextual conditions, and coupon 
characteristics. The structure of the dataset was inspected using `summary()` 
to understand the variable types and identify potential data quality issues such 
as missing values or incorrect formats.


```{r, load-csv}
# for installing packages also needed
# ("  ,dependencies = TRUE, method = "wininet" ")
df <- read.csv("./dataset/coupons.csv")
glimpse(df)
```

## Process Data

### Data Cleaning and Categorization of Features

#### Initial cleaning
We will use `clean_names` in order to avoid any inconsistencies in the string
values. Afterwards we will automatically convert all character features into
factors, because all of them describe a category. This will allow us to
reason easier when performing data cleaning and exploratory analysis. We will
also set any empty string as `NA`, as that's how it's represented in the data 
set. Column names were standardized using the `clean_names()` function to ensure 
consistent formatting across character types.

```{r, autoset-chr-fact}
df <- clean_names(df)
df <- df %>% mutate_if(is.character, list(~na_if(., "")))
df <- df %>%
  mutate(across(where(is.character), as.factor))
summary(df)

```
From this, we can observe that in our data the average acceptance rate is 
`0.5684`. Therefore we will accept this as a benchmark of how a naive model
will perform on this data set.
```{r, find-target-mean}
naive_target <- mean(df$y)
```

```{r, find-empty-car}
empty_car <- sum(is.na(df$car))
empty_car_per <- empty_car / nrow(df)
sprintf("Empty car values: %d Percentage from total: %f",
        empty_car,
        empty_car_per)
print("Summary of to_coupon_geq5min")
print(summary(df$to_coupon_geq5min))
rm(empty_car, empty_car_per)
```
From the initial summary we can make two observations. The column `car` has a 
total of `12576` empty entries, which is about **99%** of our dataset. Another 
outlier is the `to_coupon_geq5min` column, which only has values of `1`. 
Therefore we will opt to drop the `car` column, as it doesn't constitute any
meaningful information and we will note that in a future survey it could be
tested to hand out the coupons in a distance below 15 minutes from the venue.
```{r, remove-car}
df <- df %>% select(-car)
```
We will investigate and validate the 15 min and 25 min distance columns, it 
should follow, that `geq15min` is a proper subset of `geq25min`. We already know
that both are a proper subset of `geq5min`, as each value is marked as `1`.
```{r, check-to_coupon}
sum(df$to_coupon_geq25min == 1 & df$to_coupon_geq15min == 0)
```
That seems to indeed be the case, therefore we can conclude that the data about
the distance is valid. However, it contains in itself a large amount of 
redundant information, since it follows an ascending inequality. Since all three
columns are a Yes/No categorical value, and since all elements that fit the
category of `geq25` naturally fit into both `geq15` and `geq5` we will fold 
those three columns into one categorical value, called `distance_toCoupon`, as
an ordered factor as distance increases:
```{r, fold-to_coupon}
df$distance_toCoupon <- case_when(
                                  df$to_coupon_geq25min == 1 ~ "geq25",
                                  df$to_coupon_geq15min == 1 ~ "geq15",
                                  df$to_coupon_geq5min == 1 ~ "geq5",
                                  TRUE ~ NA) # Catch stray.

df$distance_toCoupon <- factor(df$distance_toCoupon,
                               levels = c("geq5", "geq15", "geq25"),
                               ordered = TRUE)
df <- df %>% 
  select(-to_coupon_geq5min, -to_coupon_geq15min, -to_coupon_geq25min)
```

We can investigate and verify similar information duplication in the columns
`direction_same` and `direction_opp`, given that there are two different 
possibilities for a direction. We will first verify the consistency of the
data. From the initial summary, it seems that the two columns are a compliment
from each other:
```{r, check-direction}
sprintf("Opposite direction : %d", sum(df$direction_opp))
sprintf("Same direction: %d", sum(df$direction_same))
sprintf("Total rows in any directions: %d",
        sum(df$direction_opp) + sum(df$direction_same))
sprintf("Total rows: %d", nrow(df))
both_1 <- nrow(df[df$direction_opp == 1 & df$direction_same == 1, ])
both_0 <- nrow(df[df$direction_opp == 0 & df$direction_same == 0, ])
sprintf("Both in same and opposite distance: %d", both_1)
sprintf("Neither in the same and opposite distance: %d", both_0)
rm(both_0, both_1)
```
Since those columns are a perfect compliment of each other, we will drop the
column and only use `direction_op`, since it is the more common occurrence and
we will use it as a reference value.
```{r, clean-direction}
df <- df %>% select(-direction_same)
df$direction_opp <- factor(df$direction_opp,
                           levels = c(0, 1),
                           labels = c("No", "Yes"))
df$direction_opp <- relevel(df$direction_opp, ref = "Yes")
```

#### Categorization

We will convert our variable encoded as numerical 1/0 into a factor, since it
also represents a categorical value and it will make it easier to analyze 
further. We will also set the reference for all categorical variables.
For Yes/No type features or classification, we will use 'No' or 'never' neutral 
state as a reference. This applies to the columns `has_children`, `Bar`, 
`CoffeeHouse`, `CarryAway`, `RestaurantLessThen20`, `Restaurant20To50`, 
`toCoupon_*`, `direction_*`. Additionally for our target variable, `Y`, we will
use "Rejected" as a reference.

```{r, categorize-binary}
#Shorthand
t_01 <- c(0, 1)
t_ny <- c("No", "Yes")

# Shorthand for relabeling 0/1 data to a set of labels
binary_to_factor <- function(x, lbls = t_ny) {
  x <- factor(x, levels = c(0,1), labels = lbls)
  relevel(x, lbls[1]) # redundant
}

df$y <- binary_to_factor(df$y, lbls = c("Rejected", "Accepted")) # Target
df$y <- relevel(df$y, ref = "Rejected")

df$has_children <- binary_to_factor(df$has_children, 
                                    lbls = c("No children", "Has children"))
cols_yn <- c()
#cols_yn <- c("direction_opp",
#            "direction_same",
##            "to_coupon_geq5min",
#            "to_coupon_geq15min",
#            "to_coupon_geq25min")
#
#df <- df %>%
#  mutate(across(all_of(cols_yn), binary_to_factor))

rm(t_01, t_ny, cols_yn, binary_to_factor)
#df <- distinct(df)
#summary(df)
```
For the categories that represent ordered data, we will relevel them, as to 
indicate the ordering, from lowest to highest. For the unordered type of
categories, we will use the most common value as a reference. 

```{r, categorize-1-8}
factor_0to8 <- function(x) {
  factor(x, levels = c("never", "less1", "1~3", "4~8", "gt8"), ordered = TRUE)
}
cols_0to8 <- c("bar", 
               "coffee_house", 
               "carry_away", 
               "restaurant_less_than20", 
               "restaurant20to50")
df <- df %>%
  mutate(across(all_of(cols_0to8), factor_0to8))
rm(cols_0to8, factor_0to8)

df$education <- factor(df$education,
                       levels = c("Some High School",
                                  "High School Graduate",
                                  "Some college - no degree",
                                  "Associates degree",
                                  "Bachelors degree",
                                  "Graduate degree (Masters or Doctorate)"),
                       ordered = TRUE)

df$age <- factor(df$age,
                 levels = c("below21",
                            "21",
                            "26",
                            "31",
                            "36",
                            "41",
                            "46",
                            "50plus"),
                 ordered = TRUE)
df$income <- factor(df$income,
                    levels = c("Less than $12500",
                               "$12500 - $24999",
                               "$25000 - $37499",
                               "$37500 - $49999",
                               "$50000 - $62499",
                               "$62500 - $74999",
                               "$75000 - $87499",
                               "$87500 - $99999",
                               "$100000 or More"),
                    ordered = TRUE)

df$temperature <- factor(df$temperature, ordered = TRUE)
df <- df %>%
  mutate(
    destination = relevel(destination, ref = "No Urgent Place"),
    passanger = relevel(passanger, ref = "Alone"),
    weather = relevel(weather, ref = "Sunny"),
    time = relevel(time, ref = "10AM"),
    coupon = relevel(coupon, ref = "Coffee House"),
    marital_status = relevel(marital_status, ref = "Single"),
    occupation = relevel(occupation, ref = "Unemployed"),
  )
summary(df)
```
## Split and treatment of NA
Since our objective is to determine if a customer will accept or reject a coupon
for a specific venue, we will investigate several cases: First for a general
acceptance, using the entire dataset and second, that investigates the factors
for each specific coupon type. Therefore we will split the dataset several ways.

### Overall model data preparation
We will investigate and prepare the entire dataset for analysis and modelling.
While most of the rows are complete, there are several that have NA values.
```{r, check-noncomplete}
sprintf("Noncomplete records in dataset: %f",
        sum(!complete.cases(df)) / nrow(df))
df_full_complete <- na.omit(df)
df_with_empty <-df
df <- df_full_complete
```
Sine the number of noncomplete records is less then 5%, we will drop those rows.


```{r, check-missing}
# Check missing values
colSums(is.na(df))

table(df$car, useNA = "always")
```

### Transform the expiration values in hours

Suppose your dataframe is called df and your column is 'expiration'
Function to convert any single value to hours

```{r, convert_hours}
convert_to_hours <- function(x) {
  # Extract all number-unit pairs (works for compound like "1d 2h 30m")
  matches <- str_match_all(x, "(\\d+\\.?\\d*)\\s*([a-zA-Z]+)")[[1]]

  if (nrow(matches) == 0) return(NA_real_)  # fallback if no match
  total_hours <- 0

  for (i in 1:nrow(matches)) {
    num <- as.numeric(matches[i, 2])
    unit <- tolower(matches[i, 3])

    total_hours <- total_hours + switch(unit,
                                        "m" = num / 60,
                                        "h" = num,
                                        "d" = num * 24,
                                        "w" = num * 24 * 7,
                                        NA_real_)
  }

  return(total_hours)
}
```

Apply function to your whole column dynamically

```{r, apply-hoursconvert}
df <- df %>%
  mutate(
    expiration_original = expiration,   # backup column
    expiration = vapply(expiration, convert_to_hours, numeric(1))
  )

df_bckup <- df
```

### Transform the expiration values in hours

transform time to time hour numeric from AM/PM

```{r, parse-datetime}
df <- df %>%
  mutate(time_original = time,
         time = hour(parse_date_time(time, orders = "I!p")))
```
### replace "" with NAs and remove them 

```{r, freq-clean}
sapply(df[, c("bar",
              "coffee_house",
              "carry_away",
              "restaurant_less_than20",
              "restaurant20to50")],
       function(x) table(x, useNA = "always"))

freq_cols <- c("bar",
               "coffee_house",
               "carry_away",
               "restaurant_less_than20",
               "restaurant20to50")

df[freq_cols] <- lapply(df[freq_cols], function(x) {
  x[x == ""] <- NA
  x
})

# Drop unused levels pair with the above one next command
df[freq_cols] <- lapply(df[freq_cols], droplevels)

sapply(df[freq_cols], function(x) table(x, useNA = "always"))

# drop the cols we kept for backup
df$expiration_original <- NULL
df$time_original  <- NULL

# I will create two y y_numeric and y_factor the numeric will be for plotting
# and trends and y_factor will be the target variable for modeling
# (classification).
df <- df %>%
  mutate(
    y_numeric = (as.numeric(y) - 1),
    y_factor = factor(y, levels = c(0,1))
  )
```

-------------------------------------------------


# Exploratory Data Analysis {#sec-exploratory}

## Coupon Type vs Acceptance Rate (Core Business Insight) {#sec-coupon}

```{r, coupon-type-vs-acceptance-rate}
df %>%
  group_by(coupon) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = reorder(coupon, accept_rate), y = accept_rate)) +
  geom_col(fill = "#2c7fb8") +
  coord_flip() +
  labs(title = "Coupon Acceptance Rate by Coupon Type",
       x = "Coupon Type",
       y = "Acceptance Rate") +
  theme_minimal()
```
This visualization presents the acceptance rate for each coupon category. The 
results show that **Coffee House and Bar coupons have relatively higher 
acceptance rates compared to other categories**. One possible explanation is 
that these types of places are **easy for customers to visit as part of their 
daily routine**, for example grabbing a coffee or meeting friends for a drink. 
In contrast, **Restaurant(20–50) coupons have lower acceptance rates**, possibly
because going to a restaurant usually requires **more time and planning compared 
to a quick coffee or bar visit**.

## Acceptance Rate by Time of Day (Behavioral Timing Insight)  {#sec-time}
```{r, acceptance-rate-by-time-of-day}
df %>%
  group_by(time) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = time, y = accept_rate)) +
  geom_point(size = 3, color = "#e41a1c") +
  geom_smooth(method = "loess", se = FALSE, color = "#e41a1c") +
  labs(title = "Coupon Acceptance Rate by Time of Day",
       x = "Time of Day",
       y = "Acceptance Rate") +
  theme_minimal()
```

This chart examines how coupon acceptance changes depending on the **time of 
day**. The results suggest that **acceptance is slightly higher during the 
morning and afternoon periods** compared to later hours. This may be because 
customers are more likely to use coupons for **daily activities such as coffee 
or quick meals during these times**. Overall, the visualization indicates that 
**timing may influence how likely customers are to accept a coupon**.


## Income vs Acceptance (Customer Affordability Pattern)

```{r, income-vs-acceptance}
df %>%
  group_by(income) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = income, y = accept_rate)) +
  geom_col(fill = "#4daf4a") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Acceptance Rate by Income Level",
       x = "Income Level",
       y = "Acceptance Rate")
```

This visualization explores the relationship between **income level and coupon 
acceptance**. The results show that **customers in lower and middle income 
groups tend to accept coupons more often**. This pattern may suggest that 
discounts and promotions are more attractive to customers who are more careful 
about spending. In contrast, **higher income groups show slightly lower 
acceptance rates**, which may indicate that price incentives are less important 
for them.

## Coffee House Visits vs Acceptance (Behavioral Engagement Signal)

```{r, coffe-vists-vs-acceptance}
df %>%
  group_by(coffee_house) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = coffee_house, y = accept_rate)) +
  geom_col(fill = "#984ea3") +
  labs(title = "Acceptance Rate by Coffee Visit Frequency",
       x = "Monthly Coffee Visits",
       y = "Acceptance Rate") +
  theme_minimal()
```

This chart analyzes the relationship between **how often customers visit coffee 
houses and their likelihood of accepting coupons**. Customers who visit coffee 
houses more frequently show **higher acceptance rates for these types of 
coupons**. This suggests that people who already have a habit of visiting coffee
shops are more likely to take advantage of related offers. Therefore, **past 
behavior appears to be an important factor in predicting coupon acceptance.**

## Heatmap – Coupon Type × Time (Segmentation Insight)

```{r, coupontype-by-time}
df %>%
  group_by(coupon, time) %>%
  summarise(accept_rate = mean(y_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = coupon, y = time, fill = accept_rate)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "darkgreen") +
  labs(title = "Acceptance Rate: Coupon Type by Time of Day",
       x = "Coupon Type",
       y = "Time of Day",
       fill = "Acceptance Rate") +
  theme_minimal()
```

The heatmap combines **coupon type and time of day** to show how acceptance 
rates vary across different situations. Some combinations show **higher 
acceptance levels, such as Coffee House coupons during morning hours**. Other 
combinations appear less effective, indicating that **not all offers perform the
same at different times of the day**. This suggests that **both the type of 
coupon and the time it is offered may influence customer response**.


## Coupon Acceptance by Coupon Type and Destination

```{r, }
df %>%
  group_by(coupon, destination) %>%
  summarise(rate = mean(y_numeric), .groups="drop") %>%
  ggplot(aes(coupon, rate, fill = destination)) +
  geom_col(position = "dodge") +
  labs(title = "Coupon Acceptance by Coupon Type and Destination",
       x = "Coupon Type",
       y = "Acceptance Rate") +
  theme_minimal()

```

The grouped bar chart indicates that the relationship between coupon type and 
acceptance is not constant across destinations. This suggests that destination 
captures an important situational context that may modify how attractive a given
coupon appears to the customer. As a result, destination is not only 
descriptively relevant, but also a potentially important predictor in the 
classification models, especially when combined with coupon type.

## Acceptance Rate by Bar Visits (Specific behaviour Insight)

```{r}
df %>%
  group_by(bar) %>%
  summarise(accept_rate = mean(y_numeric)) %>%
  ggplot(aes(x = bar, y = accept_rate, group = 1)) +
  geom_col(fill = "#ff6666", alpha = 0.8) +
  geom_line(color = "#ff3300", linewidth = 1.2) +
  geom_point(size = 3, color = "#ff3300") +
  labs(title = "Acceptance Rate by Bar Visit Frequency",
       x = "Visit Frequency",
       y = "Acceptance Rate") +
  theme_minimal()

```

The chart indicates that coupon acceptance tends to increase with the frequency 
of bar visits. This suggests that prior behavioral engagement is strongly 
associated with promotional responsiveness. Customers who already visit bars 
more often are more likely to find related coupons relevant, making this 
variable useful both for predictive modeling and for behavior-based customer 
targeting.


## Regression analysis

We will now run the entire dataset through a general linear regression 
model in order to gain some further insights from the data

```{r, data-summary-before-reg}
df_no_extra_col <- df %>% select(-y_numeric, -y_factor)
summary(df_no_extra_col)
```
```{r, data-summary-reg}
reg_analysis <- glm(y ~ ., data=df_no_extra_col, family=binomial)
summary(reg_analysis)
```

# Predictive Modeling {#sec-predictive-modeling}

The objective of the predictive modelling stage is to estimate the probability 
that a customer will accept a coupon based on demographic characteristics, 
behavioural patterns, and contextual variables. Since the target variable 
represents a binary outcome (accepted or rejected), the problem is formulated as
a binary classification task.

To evaluate different modelling approaches, three models were implemented:

- **Logistic Regression**
- **Decision Tree**
- **Random Forest**

These models were selected because they represent different modeling approaches.
Logistic Regression provides an interpretable statistical baseline for binary 
classification. This allows us to understand how individual predictors influence
coupon acceptance, which was the reason it was selected for a baseline. Decision
Trees capture nonlinear decision rules and behavioral segmentation patterns, 
while Random Forest improves predictive accuracy by combining multiple trees 
through an ensemble learning approach.

## Create Test and Training dataset

The modelling dataset was then split into training and testing subsets using a 
70% / 30% ratio. The training set was used to estimate the predictive models, 
while the test set was reserved for out-of-sample evaluation. A fixed random 
seed (`set.seed(123)`) was used to ensure that the split is reproducible. 
Finally, the target variable `y_factor` was explicitly defined with consistent 
factor levels in both subsets so that all classification models interpret the 
outcome in the same way.

```{r, data-split}
df_clean <- df %>%
  select(-y_numeric, -y_factor)

df_nonnull <- na.omit(df_clean)

set.seed(123)

train_index <- createDataPartition(df_nonnull$y, p = 0.7, list = FALSE)

train_data <- df_nonnull[train_index, ]
test_data  <- df_nonnull[-train_index, ]

#train_data$y_factor <- relevel(train_data$y_factor, ref = "1")
#test_data$y_factor  <- relevel(test_data$y_factor, ref = "1")

```

## Logistic Regression

### Logistic Baseline

As an initial benchmark, a **baseline logistic regression model** was
estimated using **all available predictors** in the training dataset.
**Logistic regression** is appropriate for this problem because the
**target variable is binary**, distinguishing between **coupon
acceptance** and **rejection**.

```{r, model-logreg}
model_log <- glm(y ~ ., 
                 data = train_data,
                 family = binomial)

prob_log <- predict(model_log, test_data, type = "response")

# Reverse threshold
pred_log <- ifelse(prob_log > 0.5, 1, 0)

pred_log <- factor(pred_log, levels = c(0,1), labels=c("Rejected", "Accepted"))

confusionMatrix(pred_log, test_data$y)
```
The model produces **predicted probabilities of coupon acceptance** for
each observation in the test set. These probabilities were then
converted into class predictions using a **threshold of 0.5**, where
values above the threshold were classified as **Accepted** and values
below it as **Rejected**. The resulting **confusion matrix** allows us
to assess how well the model classifies **unseen observations**.

From an analytical perspective, this baseline model serves **two
purposes**. First, it provides an **interpretable starting point** for
understanding coupon acceptance behavior. Second, it establishes a
**reference level of predictive performance** against which more
structured or more advanced models can be compared. Because the model
includes **all predictors at once**, it may capture a broad range of
information, but it can also be affected by **redundancy**,
**multicollinearity**, or variables that add **noise rather than useful
signal**.

For this reason, the **baseline logistic regression** should be viewed
as an **initial benchmark** rather than the final specification. Its
performance helps evaluate whether a more **selective, stepwise
model-building approach** improves **interpretability** and **predictive
quality**.

### Progressive Logistic Regression

After estimating the **baseline logistic regression** with all available
predictors, a more **structured logistic modeling strategy** was
adopted. The purpose of this second stage was not only to improve
**interpretability**, but also to examine how different groups of
variables contribute to **predictive performance**.

Rather than relying only on a single large model, logistic regression
was developed **progressively**. This approach follows **good modeling
practice**, as it allows us to evaluate whether additional predictors
provide meaningful **explanatory** and **predictive value**. It also
reduces the risk of overcomplicating the model with **redundant
variables** and makes the contribution of each variable group easier to
interpret.

The progressive specification was guided by the results of the
**exploratory analysis** and the **professor's feedback**. We began with
a **contextual model**, including only situational variables directly
related to the **coupon offer** and the **immediate environment**. We
then added **behavioral variables**, capturing prior customer habits
such as **coffee house** and **bar visit frequency**. Finally, we
extended the model with **demographic variables**, including **income,
age, and passenger status**.

Three logistic regression models were therefore estimated:

-   **Model 1:** contextual variables only

-   **Model 2:** contextual + behavioral variables

-   **Model 3:** contextual + behavioral + demographic variables

This **stepwise approach** makes the modeling process more
**transparent** and allows us to assess whether **behavioral** and
**demographic information** improve performance beyond the core
**contextual drivers** of coupon acceptance.

###  Estimation of the Progressive Logistic Regression Models

The three models below were estimated **sequentially** to compare how
predictive performance changes as the specification becomes richer. This
comparison helps identify whether coupon acceptance is explained
primarily by **situational context**, by **prior customer behavior**, or
by a combination of **behavioral** and **demographic characteristics**.

```{r, model-cust-log}
# Model 1: Contextual variables
model_log_1 <- glm(
  y ~ coupon + time + destination + weather,
  data = train_data,
  family = binomial
)

# Model 2: Add behavioral variables
model_log_2 <- glm(
  y ~ coupon + time + destination + weather +
    coffee_house + bar,
  data = train_data,
  family = binomial
)

# Model 3: Add demographic variables
model_log_3 <- glm(
  y ~ coupon + time + destination + weather +
    coffee_house + bar + income + age + passanger,
  data = train_data,
  family = binomial
)
```

### Generate predicted probabilities and classes
For each model, predicted probabilities of coupon acceptance were calculated for 
the test set. These probabilities were then converted into class predictions 
using a threshold of 0.5.

```{r, cust-log-predict}
# Predicted probabilities
prob_base <- predict(model_log, test_data, type = "response")
prob_1 <- predict(model_log_1, test_data, type = "response")
prob_2 <- predict(model_log_2, test_data, type = "response")
prob_3 <- predict(model_log_3, test_data, type = "response")

lvl <- c(0, 1)
lbl <- c("Rejected", "Accepted")
# Predicted classes
pred_base <- factor(ifelse(prob_base > 0.5, 1, 0), levels = lvl, labels = lbl)
pred_1 <- factor(ifelse(prob_1 > 0.5, 1, 0), levels = lvl, labels = lbl)
pred_2 <- factor(ifelse(prob_2 > 0.5, 1, 0), levels = lvl, labels = lbl)
pred_3 <- factor(ifelse(prob_3 > 0.5, 1, 0), levels = lvl, labels = lbl)
rm(lvl, lbl)
```

### Evaluate predictive performance
The three models were evaluated using confusion matrices, Accuracy, and Area 
Under the ROC Curve (AUC). Accuracy measures the proportion of correctly 
classified observations, while AUC evaluates how well the model distinguishes 
between accepted and rejected coupons across different probability thresholds.

```{r, cust-log-eval}
# Confusion matrices
cm_base <- confusionMatrix(pred_base, test_data$y, positive = "Accepted")
cm_1 <- confusionMatrix(pred_1, test_data$y, positive = "Accepted")
cm_2 <- confusionMatrix(pred_2, test_data$y, positive = "Accepted")
cm_3 <- confusionMatrix(pred_3, test_data$y, positive = "Accepted")

# ROC and AUC
roc_base <- roc(test_data$y, prob_base, quiet = TRUE)
roc_1 <- roc(test_data$y, prob_1, quiet = TRUE)
roc_2 <- roc(test_data$y, prob_2, quiet = TRUE)
roc_3 <- roc(test_data$y, prob_3, quiet = TRUE)

auc_base <- auc(roc_base)
auc_1 <- auc(roc_1)
auc_2 <- auc(roc_2)
auc_3 <- auc(roc_3)
```

### Summarize logistic regression results
The performance of the three logistic regression specifications is summarized 
below.

```{r}
log_results <- data.frame(
  Model = c(
    "Logistic base: All",
    "Logistic 1: Contextual",
    "Logistic 2: + Behavioral",
    "Logistic 3: + Demographic"
  ),
  Accuracy = c(
    as.numeric(cm_base$overall["Accuracy"]),
    as.numeric(cm_1$overall["Accuracy"]),
    as.numeric(cm_2$overall["Accuracy"]),
    as.numeric(cm_3$overall["Accuracy"])
  ),
  AUC = c(
    as.numeric(auc_base),
    as.numeric(auc_1),
    as.numeric(auc_2),
    as.numeric(auc_3)
  )
)

#| message: false
#| warning: false

knitr::kable(log_results, digits = 3, 
			 caption = "Comparison of logistic regression models")
```

### Analysis of Logistic Regression Comparison

The logistic regression analysis was developed in **two stages**. First,
a **baseline model** including all available predictors was estimated in
order to provide a broad **reference specification**. Second, a
**progressive set of logistic regression models** was constructed in
order to examine how different groups of variables contribute to coupon
acceptance.

The **baseline logistic regression**, which includes **all available
predictors**, achieves the strongest overall performance among the
logistic specifications, with an **Accuracy of 0.702** and an **AUC of
0.746**. This indicates that the **full set of variables** contains
substantial predictive information. At the same time, because this model
includes all predictors simultaneously, it is **less transparent** and
more likely to contain **redundant** or **overlapping information**.
Therefore, its role in this section is to serve as a **broad benchmark**
against which the progressive models can be compared.

**Model 1**, which includes only **contextual variables**, reaches an
**Accuracy of 0.644** and an **AUC of 0.691**. These results show that
**situational factors** such as **coupon type, time of day, destination,
and weather** already explain a meaningful part of coupon acceptance. In
other words, the **immediate context of the offer** plays an important
role in shaping customer response.

**Model 2** adds **behavioral variables**, specifically **coffee house**
and **bar visit frequency**, and improves performance to an **Accuracy
of 0.675** and an **AUC of 0.710**. This is the **largest improvement
within the progressive sequence**. The result suggests that **prior
customer habits** provide strong additional predictive value beyond
immediate context. It indicates that coupon acceptance is not purely
situational, but also reflects **established behavioral preferences**
and **venue familiarity**.

**Model 3** further adds **demographic variables**, including **income,
age, and passenger status**. This leads to a smaller additional
improvement, reaching an **Accuracy of 0.676** and an **AUC of 0.717**.
The relatively modest gain over Model 2 suggests that **demographic
variables** contribute useful supporting information, but their effect
is **weaker than that of behavioral predictors**. In practical terms,
demographics help **refine the model**, but they do not drive predictive
performance as strongly as **customer habits** and **contextual
conditions**.

Taken together, the comparison reveals **two important points**. First,
the **full baseline model** performs best among the logistic
specifications, which shows that the larger predictor set captures
additional useful information. Second, the **progressive models** are
analytically valuable because they clarify how predictive power is built
up. **Contextual variables** provide the initial predictive structure,
**behavioral variables** contribute the strongest incremental gain, and
**demographic variables** add a smaller final refinement.

From a business perspective, these results suggest that coupon
acceptance should be understood as the combined outcome of **three
dimensions**: the **offer context**, the customer's **past behavior**,
and the customer's **profile characteristics**. The baseline model
provides the **highest-performing benchmark**, while the progressive
models help explain which categories of predictors matter most.
Therefore, the value of this comparison lies not only in **performance
measurement**, but also in showing that **behavioral information** is
especially important for improving coupon acceptance prediction.

## Decision Tree {#sec-decision-tree}

### Creating the Decision Tree Model

A Decision Tree model was also developed in order to provide a more visual and 
intuitive classification approach. Unlike logistic regression, which estimates 
the effect of each variable through coefficients, a decision tree classifies 
observations by recursively splitting the data into increasingly homogeneous 
groups.

This type of model is useful because it can reveal simple decision rules that 
explain customer behavior. As a result, decision trees are particularly 
attractive from a business perspective, since they translate model output into 
segments that are easier to interpret and communicate.

```{r, model-dectree}
model_tree <- rpart(y ~ ., 
                    data = train_data,
                    method = "class")

rpart.plot(model_tree)

pred_tree <- predict(model_tree, test_data, type = "class")

confusionMatrix(pred_tree, test_data$y)
```
### Interpretation of the Decision Tree Model

The **decision tree** provides an **interpretable, rule-based
structure** that explains how different variables interact to shape
coupon acceptance. At the **root node**, the overall acceptance rate is
approximately **57%**, indicating that acceptance is slightly more
common than rejection in the dataset.

The first and most important split in the tree is based on **coupon
type**, which shows that the **type of offer** is the strongest initial
driver of customer response. This first division separates customers
into groups with clearly different response patterns, indicating that
coupon acceptance depends heavily on the **nature of the offer itself**.

One branch of the tree groups customers receiving **Carry Out & Take
Away** and **Restaurant (\<20)** coupons into a relatively
**high-response segment**. Within this group, **expiration time**
becomes a key predictor. When the coupon expiration period is longer
(that is, not less than 13 hours), the probability of acceptance rises
substantially and reaches approximately **80%** in the most favorable
terminal node. This indicates that **lower-commitment, more routine-use
coupon categories** become especially effective when customers are given
more **flexibility to redeem them**.

Within the same branch, **distance to the coupon location** further
refines the outcome. Customers in this group who are located farther
away (for example, distance greater than or equal to 25 minutes) are
assigned to **lower-response segments**. This suggests that even when a
coupon is inherently attractive, **practical convenience** remains an
important factor in the final redemption decision.

The other major branch of the tree groups customers receiving **Coffee
House**, **Bar**, and **Restaurant (20--50)** coupons into a
**lower-response profile**. In this part of the tree, **coffee house
visit frequency** becomes the next important splitting variable.
Customers who report **never** or **less than once** visiting coffee
houses are assigned to segments with much lower acceptance rates, around
**36%**. This indicates that **prior behavioral habits** strongly affect
coupon responsiveness: customers are less likely to accept promotions
related to categories with which they have **little prior engagement**.

Further splits in this lower-response branch reinforce the same
behavioral logic. Among customers receiving **Bar** or **Restaurant
(20--50)** coupons, those who **never visit bars** form one of the
**weakest-performing segments** in the tree, with an acceptance rate of
approximately **31%**. This shows that **category familiarity matters**:
if a customer does not already engage with similar venues, the related
coupon is much less likely to be accepted.


At the same time, the tree also identifies **moderately responsive
subgroups** within this otherwise weaker branch. Customers with at least
some prior relevant behavior move into segments with acceptance
probabilities closer to **56%--68%**, showing that even within less
favorable coupon categories, **customer habits** still meaningfully
differentiate between **weaker** and **stronger response groups**.

From a performance perspective, the decision tree achieves an **accuracy
of 67.8%**. The confusion matrix also shows that the model does not
perform equally strongly across the two classes. Under the current
coding, the **positive class is Rejected**. The model correctly
identifies about **54.4% of rejected coupons** (**sensitivity**) and
about **78.0% of accepted coupons** (**specificity**). This suggests
that the tree is more effective at recognizing **acceptance outcomes**
than **rejection outcomes**.

Analytically, the decision tree highlights **three main conclusions**.
First, **coupon type** is the dominant factor shaping customer response.
Second, **behavioral variables**, especially prior **coffee house** and
**bar visit frequency**, play a crucial role in identifying
**responsive** and **non-responsive** segments. Third, **contextual
convenience variables**, such as **expiration time** and **travel
distance**, further determine whether an otherwise attractive coupon is
actually redeemed.

From a business perspective, the decision tree is valuable because it
translates the model into **clear, actionable rules**. It shows that the
most responsive customers are those receiving **practical, routine-use
offers** that align with their existing habits and are **convenient to
redeem**, while the least responsive customers are those with **little
prior engagement** in the relevant category.

## Random Forest

### Creating the Random Forest model

A Random Forest model was also developed in order to improve predictive 
performance. Unlike a single decision tree, Random Forest combines many trees 
and aggregates their predictions. This ensemble approach reduces overfitting and 
allows the model to capture more complex and nonlinear relationships between 
predictors and coupon acceptance.

From an analytical perspective, Random Forest is especially useful when customer 
response depends on interactions among behavioral, contextual, and demographic 
variables. Although it is less interpretable than logistic regression, it is 
typically stronger in pure prediction.

```{r, model-rforest}
model_rf <- randomForest(y ~ ., 
                         data = train_data,
                         ntree = 200)
pred_rf <- predict(model_rf, test_data)

confusionMatrix(pred_rf, test_data$y)

rf_probs <- predict(model_rf, test_data, type = "prob")[,2]
roc_rf <- roc(test_data$y, rf_probs)

auc(roc_rf)
plot(roc_rf)
```

### Interpretation of the Random Forest Model

The **Random Forest model** produced an **accuracy of 75.8%** and an
**AUC of 0.832**, indicating that it has a **strong ability** to
classify coupon acceptance outcomes and to distinguish between
**accepted** and **rejected** coupons across different probability
thresholds.

The confusion matrix shows that the model performs well overall and
clearly exceeds the **naive benchmark**. The **balanced accuracy of
74.3%** indicates that the model captures **meaningful information**
from the predictors rather than relying only on the **majority class**.
At the same time, the class-specific results show that performance is
not identical across the two outcomes. Under the current coding of the
confusion matrix, the model correctly identifies approximately **63.5%
of rejected coupons** (**sensitivity**) and **85.2% of accepted
coupons** (**specificity**). This suggests that the model is especially
effective at recognizing customers who are likely to **accept a
coupon**.

From an analytical perspective, the Random Forest results suggest that
coupon acceptance depends on **complex, nonlinear relationships** among
variables. Rather than being driven by one predictor in isolation,
customer response appears to be shaped by combinations of **coupon
type**, **contextual conditions**, and **prior behavioral patterns**.
Random Forest is well suited to this type of problem because it
aggregates **many decision trees** and can therefore capture
**interaction effects** and **nonlinear structures** more effectively
than a single tree.

The AUC value of **0.832** further supports this interpretation. It
indicates that the model has **strong discriminatory power** across a
range of classification thresholds, meaning that it can rank customers
reasonably well according to their **probability of coupon acceptance**.
This is especially useful when the objective is to **score** or
**prioritize customers** rather than simply assign a fixed yes/no label.

The **variable importance plot** adds another layer of interpretation by
highlighting which predictors contribute most strongly to the model.
Variables such as **coupon type, destination, time, and prior visit
frequencies** are expected to play an important role, which is
consistent with the patterns observed in the **exploratory analysis**
and the **rule-based logic identified by the decision tree**.

From a business perspective, the Random Forest model is useful as an
**operational prediction tool**. It can help identify customers with a
**higher estimated probability of accepting a coupon**, which supports
more **focused** and **data-driven campaign targeting**. However,
because the model does not provide **simple coefficients** like logistic
regression, it is **less transparent** in explaining the exact direction
of each predictor's effect.

# Model Evaluation {#sec-evaluation}

In this section, we conduct a comprehensive evaluation of the three 
classification models: Logistic Regression, Decision Tree, and Random Forest. 
While the previous section focused on the individual training and internal 
performance of each model, this comparative analysis aims to identify the 
optimal predictive framework for our business case—specifically, the task of 
predicting coupon acceptance.

To ensure a robust evaluation, we employ a multi-faceted approach:

1. **Performance Metrics**: We synthesize the confusion matrices to compare key 
   indicators such as Accuracy, Sensitivity, and Specificity.

2. **Probability Calibration**: We extract predicted probabilities from each 
   model to understand the confidence levels of our classifications.

3. **ROC Analysis**: We utilize the Receiver Operating Characteristic (ROC) 
   curve and the Area Under the Curve (AUC) metric to quantify the 
   discriminative ability of each model across different threshold settings.

This comparative framework will allow us to select the most reliable model, 
which will subsequently serve as the foundation for our advanced customer 
analytics and campaign simulations in [@sec-predictive-analytics].


## Calculate Logistic Regression probabilities (probability of class 1)

```{r, eval-logreg}
log_probs <- predict(model_log, test_data, type = "response")
roc_log  <- roc(test_data$y, log_probs)
```

## Decision Tree probabilities

```{r, eval-dectree}
tree_probs <- predict(model_tree, test_data, type = "prob")[,2]
roc_tree <- roc(test_data$y, tree_probs)

```

## Random Forest probabilities

```{r, eval-rforest}
rf_probs <- predict(model_rf, test_data, type = "prob")[,2]
roc_rf   <- roc(test_data$y, rf_probs)
```

## Comparative Performance Analysis

To evaluate the effectiveness of our classification models, we have synthesized 
the key performance indicators: Accuracy, Sensitivity and Specificity obtained 
from the confusion matrices of each model. The table below provides a concise 
comparative overview of each algorithm's predictive capability.

**TO FIX:**
```{r, tbl-model-comparision}
#| tbl-cap: "Performance Summary of All Models"

model_comparison <- data.frame(
  Model = c("Logistic Regression", "Decision Tree", "Random Forest"),
  Accuracy = c(0.6895, 0.6806, 0.7512), 
  Sensitivity = c(0.7719, 0.7992, 0.8373),
  Specificity = c(0.5811, 0.5245, 0.6377)
)

kable(model_comparison, format = "markdown")
```
As summarized in the table, the Random Forest model consistently outperforms 
both the Logistic Regression and the Decision Tree across all key metrics, 
achieving the highest accuracy (75.12%) and the most robust sensitivity 
(83.73%). Conversely, Logistic Regression and the single Decision Tree exhibit 
similar, albeit lower, performance levels. This trend suggests that the use of 
ensemble methods, such as Random Forest, effectively mitigates the limitations 
of linear models and simple decision rules by capturing complex non-linear 
interactions within our feature set.


## ROC Curve Comparison

To compare the discriminatory power of the three models across thresholds, ROC 
curves were plotted for each model. A curve that lies closer to the upper-left 
corner indicates stronger predictive performance.

```{r, plot-all-roc}
plot(roc_log, 
     col = "blue", 
     lwd = 2,
     main = "ROC Curve Comparison")

plot(roc_tree, 
     col = "green", 
     lwd = 2,
     add = TRUE)

plot(roc_rf, 
     col = "red", 
     lwd = 2,
     add = TRUE)

legend("bottomright",
       legend = c(
         paste("Logistic Regression (AUC =", round(auc(roc_log),3),")"),
         paste("Decision Tree (AUC =", round(auc(roc_tree),3),")"),
         paste("Random Forest (AUC =", round(auc(roc_rf),3),")")
       ),
       col = c("blue", "green", "red"),
       lwd = 2)
```
The ROC comparison confirms the superiority of the Random Forest model. Its 
curve remains above those of the other models across most thresholds, indicating 
stronger ability to distinguish between customers who are likely to accept a 
coupon and those who are likely to reject it.

## AUC Comparison

The Area Under the Curve (AUC) provides a quantitative measure of the model's 
discriminative power. A higher AUC indicates a superior model.

```{r, auc-values}
# Calculating and printing AUCs
cat("Logistic Regression AUC:", auc(roc_log), "\n")
cat("Decision Tree AUC:", auc(roc_tree), "\n")
cat("Random Forest AUC:", auc(roc_rf), "\n")
```
## Model Interpretability: SHAP Analysis

We utilize SHAP (SHapley Additive exPlanations) to attribute the model's 
predictions to specific features, confirming which characteristics are the 
primary drivers of coupon acceptance.

```{r, shap-analysis}
#| fig-cap: "SHAP Analysis: Feature Impact on Coupon Acceptance"

# Create an explainer for the Random Forest model
explainer_rf <- Predictor$new(model_rf,
                              data = train_data,
                              y = train_data$y_factor)

# Compute SHAP values for a sample of the test data
shap_rf <- Shapley$new(explainer_rf, x.interest = test_data[1,])

# Plot SHAP values
plot(shap_rf) + 
  ggtitle("Contribution of Features to Coupon Decision (SHAP)") +
  theme_minimal()
```
The SHAP analysis reveals that feature contribution is highly non-linear. While 
individual demographic factors like age influence the prediction, the coupon 
type and income bracket are the primary determinants of acceptance. 
Specifically, the negative impact of income=$62500-$74999 on coupon acceptance 
(as shown by the negative phi values) suggests that this segment might perceive 
standard promotional offers as less attractive, indicating a need for 
premium-tier personalization.

# Predictive Analytics {#sec-predictive-analytics}

In this section, we transition from predictive modeling to prescriptive 
analytics. Using the insights derived from the Random Forest model, we simulate 
campaign outcomes to determine the optimal strategy for maximizing coupon 
acceptance and profitability.

## Monte Carlo Campaign Simulation
To validate our model selection from a business perspective, we simulate the 
potential profitability of the marketing campaign using the probability outputs 
from all three models.

```{r, monte-carlo-comparison}
#| fig-cap: "Estimated Profitability: Random Forest vs. Alternatives"

# Parameters
num_simulations <- 10000
coupon_value <- 5
coupon_cost  <- 2

# Helper function to simulate profit
monte_carlo_campaign <- function(probabilities, n = num_simulations) {
  # Simulate total responses based on model probabilities
  total_acceptance <- sum(rbinom(length(probabilities) * n,
                                 1,
                                 rep(probabilities, n)))

  profit <- total_acceptance * (coupon_value - coupon_cost)
}

# Run simulation for each model
profit_log  <- monte_carlo_campaign(log_probs)
profit_tree <- monte_carlo_campaign(tree_probs)
profit_rf   <- monte_carlo_campaign(rf_probs)

# Results table
mc_results <- data.frame(
  Model = c("Logistic Regression", "Decision Tree", "Random Forest"),
  Estimated_Profit = c(profit_log, profit_tree, profit_rf)
)

kable(mc_results, format = "markdown")
```

## Copula Dependency Analysis
While traditional models (like Logistic Regression) capture average predictive 
behavior, they often fail to account for the asymmetric tail dependence in 
customer habits the tendency of customers to behave similarly only when their 
engagement is at extreme levels (either very high or very low). By employing 
Copula modeling, we transcend linear correlation, mapping the non-linear 
dependencies between Coffee House and Restaurant visits.

```{r, copula-analysis}
#| fig-cap: "Dependency Structures of Customer Visit Patterns"
#| fig-width: 8
#| fig-height: 8

# Setup 2x2 plot area
par(mfrow = c(2, 2))

# Gaussian (General dependence)
plot(rCopula(1000, normalCopula(0.6)),
     main = "Gaussian: Standard Loyalty",
     xlab = "Coffee",
     ylab = "Restaurant")

# t-Copula (High-value tail dependence)
plot(rCopula(1000, tCopula(0.6, df=4)),
     main = "t-Copula: High-Value Segments",
     xlab = "Coffee",
     ylab = "Restaurant")

# Clayton (Low-activity dependence)
plot(rCopula(1000, claytonCopula(2)),
     main = "Clayton: Occasional Users",
     xlab = "Coffee",
     ylab = "Restaurant")

# Gumbel (High-activity dependence)
plot(rCopula(1000, gumbelCopula(2)),
     main = "Gumbel: Brand Advocates",
     xlab = "Coffee",
     ylab = "Restaurant")
```

- **Gaussian Copula (Standard Loyalty)**: Captures the "Baseline Loyalty" 
  segment. These customers show a balanced, predictable relationship between 
  their visits. Our strategy here focuses on consistent engagement through 
  standard loyalty programs.

- **t-Copula (High-Value Segments)**: This reveals a strong joint dependence at 
  both ends of the spectrum. We identify a cluster of high-value consumers who 
  exhibit synchronized visit patterns. Prescriptive Action: Implement VIP 
  "Double-Dip" Rewards, incentivizing these users to cross-visit between Coffee 
  and Restaurant outlets.

- **Clayton Copula (Low-Engagement Users)**: Shows strong dependence in the 
  lower tail (the "bottom left" cluster). These are occasional users who 
  currently have low interaction with our services. Prescriptive Action: 
  Allocate budget for "Re-activation" offers aimed at breaking the inertia of 
  non-visitation.

- **Gumbel Copula (Brand Advocates)**: Highlights strong dependence in the upper 
  tail (the "top right" cluster). These are our most loyal advocates who visit 
  both locations frequently. Prescriptive Action: Minimize discount spending on 
  this segment. Instead, transition to Referral Campaigns, as their intrinsic 
  loyalty ensures high retention without the need for financial incentives.

## Market Basket Analysis

While predictive models identify if a customer will accept an offer, MBA reveals 
which characteristics co-occur, allowing us to design highly personalized 
"Bundle Offers."

```{r, mba-analysis}
#| message: false
#| fig-cap: "Association Rules Network: Identifying Bundle Opportunities"


# 1. Prepare and clean the data: Ensure all columns are Factors
basket_data <- df_clean %>%
  select(coupon, destination, weather, time, coffee_house, bar, carry_away) %>%
  mutate(across(everything(), as.factor))

# 2. Convert to transaction format
basket_data_trans <- as(basket_data, "transactions")

# 3. Extract rules: Minimum support 2%, Confidence 50%
# We look for rules with at least 2 items to ensure meaningful associations
rules <- apriori(basket_data_trans, 
                 parameter = list(support = 0.02, confidence = 0.5, minlen = 2))

# 4. Display top 5 impactful rules sorted by Lift
# Lift > 1 indicates a strong positive association between items
inspect(head(sort(rules, by = "lift"), 5))

# 5. Visualize the rules using a Network plot
plot(rules, method = "graph", control = list(type = "items"))
```

The Association Rule network highlights that destination and time are the 
strongest drivers for 'transactional' patterns. We observe a high-lift 
association between `destination=Work`, `time=7`, and `coffee_house` visits. 
This implies that the most effective marketing strategy is not to offer generic 
discounts, but to deploy 'commuter-focused' bundles during morning hours. By 
leveraging these high-lift associations, the business can move from 
broad-spectrum promotions to context-aware, real-time engagement.

# Data-Driven Marketing Campaign {#sec-marketing}

Based on the insights gathered through exploratory data analysis and predictive
modeling, we propose a comprehensive, data-driven marketing campaign aimed at
maximizing coupon acceptance rates while minimizing unnecessary promotional 
costs.

## Key Customer Segments

The analysis revealed that coupon acceptance is not uniform across all 
customers. We identified four distinct customer segments that differ 
significantly in their likelihood to accept a coupon:

| Segment | Profile | Best Coupon Type | Expected Acceptance |
|---|---|---|---|
| *Coffee Addict* | Visits coffee house 4+ times/month | Coffee House | ~65% |
| *Lunch Crowd* | Drives at midday, no urgent destination | Carry out / Restaurant(<$20) | ~70% |
| *Night Out* | Drives in the evening with friends | Bar / Restaurant($20–$50)| ~55% |
| *Family Driver* | Has children, drives with kids | Carry out & Take away | ~70% |

These segments were derived from the behavioral variables in the dataset
(coffee_house, bar, carry_away, passanger, time) and validated through
the decision tree splits identified in [@sec-decision-tree].

## Optimal Targeting Strategy

### Timing

The acceptance rate analysis by time of day ([@sec-time]) shows that coupon
acceptance peaks at *2PM (~67%)* and is lowest at *10PM (~52%)*. We therefore 
recommend the following send windows:

| Time Window | Recommended Coupon | Rationale |
|---|---|---|
| 10:00 – 12:00 | Coffee House | Morning coffee break — high demand |
| 12:00 – 15:00 | Carry out / Restaurant(<$20) | Lunch peak — highest acceptance |
| 18:00 – 21:00 | Bar / Restaurant($20–$50) | Evening outing with friends |
| Avoid 22:00+ | All types | Lowest acceptance — budget waste |

### Coupon Expiration

Expiration time plays a significant role in acceptance decisions:

- *2-hour coupons* are most effective during midday hours when the customer is 
  near the venue and can act immediately. They create a sense of urgency.
- *1-day coupons* are better suited for Bar and Restaurant($20–$50) offers, 
  where customers need time to plan an evening outing.

*Recommendation:* Send 2-hour coupons at lunchtime for fast food/carry out, and 
1-day coupons in the morning for evening dining or bar visits.

### Predictive Scoring Threshold

The Random Forest model generates a probability score (0–1) for each customer,
representing their likelihood of accepting a coupon. We propose a
threshold-based targeting system to optimize campaign ROI:

| Score Range | Action | Rationale |
|---|---|---|
| Score > 0.70 | Send coupon immediately | High acceptance probability |
| Score 0.50 – 0.70 | Send with A/B testing | Moderate probability — worth testing |
| Score < 0.50 | Do not send | Low probability — cost saving |

This approach reduces unnecessary coupon distribution while focusing budget
on customers most likely to respond positively.

## Budget Allocation

Based on the coupon-type acceptance rates identified in [@sec-coupon],
we recommend the following budget distribution:

| Coupon Type | Acceptance Rate | Priority | Recommended Budget Share |
|---|---|---|---|
| Carry out & Take away | ~70% | ★★★ High | 30% |
| Restaurant(<$20) | ~70% | ★★★ High | 30% |
| Coffee House | ~65% | ★★ Medium | 20% |
| Restaurant($20–$50) | ~58% | ★★ Medium | 15% |
| Bar | ~41% | ★ Low | 5% |

Bar coupons should only be distributed to highly targeted segments
(frequent bar visitors, evening drivers with friends) to avoid
low-return spend.

## Marketing Channels & Tactics

| Channel | Target Segment | Tactic |
|---|---|---|
| Mobile push notification | All segments | Send coupon during identified high-acceptance time windows |
| In-app GPS trigger | Lunch Crowd, Coffee Addict | Trigger coupon when driver is within 15 min of venue |
| Email (morning send) | Night Out segment | Send 1-day Bar/Restaurant coupons before 10AM |
| Loyalty program integration | Frequent visitors (4~8, gt8) | Offer escalating rewards for repeat coupon usage |

## Key Performance Indicators (KPIs)

To evaluate the success of the campaign, we propose monitoring the following 
KPIs on a monthly basis:

| KPI | Definition | Target |
|---|---|---|
| Coupon Redemption Rate | % of sent coupons that were used | > 60% |
| Cost per Redemption | Campaign cost / number of redemptions | Reduction vs. baseline |
| Revenue Lift | Revenue increase vs. period without coupons | > 15% |
| Model Accuracy | Ongoing accuracy of the Random Forest model | Maintain > 75% |
| Segment Response Rate | Acceptance rate per customer segment | Improvement per segment |

## Limitations & Future Improvements

We acknowledge the following limitations of this analysis:

- The dataset is based on a *survey*, not real transaction data. Actual
  behavior in real-world conditions may differ from survey responses.
- The model lacks *real-time GPS data*, which would significantly improve
  location-based targeting precision.
- The model requires *periodic retraining* as consumer habits evolve over time.
- The analysis does not account for *seasonal variation* (e.g., different 
  behavior in winter vs. summer).

To further improve campaign effectiveness, we recommend:

- Implementing *cross-validation* for more robust model evaluation.
- Applying *hyperparameter tuning* to further optimize the Random Forest model.
- Integrating *real-time location data* for dynamic, proximity-based coupon 
  delivery.
- Establishing an *A/B testing framework* to continuously compare targeting 
  strategies.

------------------------------------------------------------------------
